---
title: "Sea surface temperature German North Sea in R"
description: |
  Create a map with sst information from Copernicus in R.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-03-15
preview: blog46_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---



# Intro

Sea surface temperatures gives us information on Celsius around an area.

This post includes:  
- Download raster data   
- Read and subset raster data  
- Plot raster data  

# Data

Download raster data

Copernicus Marine Data Store

To download [visit webpage](https://data.marine.copernicus.eu/products)  
You need to register first  

Go to:  
- Click on the icon of download  
- [Click on download](https://help.marine.copernicus.eu/en/articles/4469993-how-to-download-copernicus-marine-products#h_ec9344c750)  
- Here I selected Sea water potential temperaturethetao [°C]  
- Add your coordinates here I use 1 to 10 and 50 to 60    
- Add you time scale here I use April and May 2018  
- I select the depths form 0 to 60 because I am using birds data  


Read raster data from a file

```{r, include=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2024-10-24-copernicus-sst")
list.files(this_folder)
```

```{r, include=FALSE}
this_file<-paste0(this_folder,"/cmems_mod_glo_phy_my_0.083deg-climatology_P1M-m_1729785319560.nc")
this_file
```

Load the [terra package](https://pjbartlein.github.io/REarthSysSci/netcdf_terra.html) for reading raster data

```{r}
library(terra)
```

The function **rast** helps to read raster data - replacing package raster

```{r}
SST_file<-rast(this_file)
```

Change to data frame

```{r}
SST_dataframe <- as.data.frame(SST_file, xy = TRUE)
```

# Subset raster data

Load the package **tidyverse** 

```{r}
library(tidyverse)
```

Use **filter** to subset your data to your geographical area

```{r}
SST_sub <-SST_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)
```

## SST info

Data was collected almost every meter

Therefore there are many columns with data per depth

```{r}
ncol(SST_sub)
```
Keep columns with the depth area

```{r}
SST_depth<-SST_sub %>%
  select(starts_with("thetao"))
```

The SST data is collected almost for every meter

```{r}
Depths<-colnames(SST_depth)
head(Depths)
```

```{r}
SST_depth_perloc<-SST_depth %>%
  rowwise()%>%
  mutate(mean_SST = mean(c_across(where(is.numeric)),na.rm=TRUE),
         min_SST = min(c_across(where(is.numeric)),na.rm=TRUE),
         max_SST = max(c_across(where(is.numeric))),na.rm=TRUE)%>%
  relocate(mean_SST,min_SST,max_SST)
```

## Check

```{r}
first_column<-SST_depth[1,]

long_values<-first_column %>% 
  pivot_longer(
    cols = 1:228, 
    names_to = "type",
    values_to = "value"
)

arrange_values<-arrange(long_values, desc(value))

head(arrange_values)
```


## Mean values

```{r}
SST_sub$SST<-SST_depth_perloc$mean_SST
```

```{r}
mean(SST_sub$SST)
```

# Plot

Load the package **ggplot2** 

```{r}
library(ggplot2)
```

Plot your data using **geom_raster**

Select [palette](https://ggplot2-book.org/scales-colour)

```{r, fig.width=6, fig.height=4}
ggplot() +
  geom_raster(data = SST_sub , aes(x = x, y = y, fill = SST)) +
  scale_fill_distiller(palette = "YlOrBr")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```

Add land for reference

Load the package **sf**

```{r}
library(sf)
```

Use the function **st_transform** to convert to the same CRS

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```

Add the land to the plot using **geom_sf**

```{r, fig.width=6, fig.height=6}
Plot_sst<-ggplot() +
  geom_tile(data = SST_sub , aes(x = x, y = y, fill = SST)) +
  scale_fill_distiller(palette = "YlOrBr")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  theme_void()+
  theme(legend.position='left')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_sst+ ggtitle('Sea surface temperature from 0 to 60 m depth \n in the German North Sea in \n spring (April and May montly composites)')
```



```{r, echo=FALSE, eval=FALSE}
ggsave(Plot_sst,
       filename = paste0(this_folder,"/sst_GermanNorthSea.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

- [Package GermanNorthSea](https://github.com/MiriamLL/GermanNorthSea)
- [Package terra](https://rspatial.github.io/terra/) 
- [Package sf](https://r-spatial.github.io/sf/)

- [Examples](https://marine.copernicus.eu/services/user-learning-services/using-r-work-copernicus-marine-data)
- [How to download copernicus marine products](https://help.marine.copernicus.eu/en/articles/4469993-how-to-download-copernicus-marine-products#h_ec9344c750)


Comparing with other studies seems to fit

SST [from 2004](https://www.researchgate.net/publication/222406095_Defining_habitats_suitable_for_larval_fish_in_the_German_Bight_southern_North_Sea_An_IBM_approach_using_spatially-_and_temporally-resolved_size-structured_prey_fields)


# End of document