---
title: "tmap"
description: |
  Create an interactive map to explore your data. 
author:
  - name: Miriam Lerma
    url: {}
date: 2024-12-04
preview: blog40_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - tmap
  - English
  - Y2024
---

# Intro

This post is to overview your locations in a interactive map.

# Data 

For the exercises, test data is from masked boobies. <br>
To access the data you have to install the package **sula**: devtools::install_github("MiriamLL/sula")

```{r}
library(sula)
```

```{r}
Data_1original<-GPS_preparado 
```

## Transform

```{r}
library(tidyverse)
```

```{r}
Data_2fortify<-fortify(Data_1original)
```

Make sure your lat and lon are numerical

```{r}
Data_2fortify$lat<-as.numeric(Data_2fortify$Latitude)
Data_2fortify$lon<-as.numeric(Data_2fortify$Longitude)
```

Transform to spatial

```{r}
library(sf)
```
Select which CRS you will like to use

```{r}
Data_3spatial <- st_as_sf(Data_2fortify, coords = c("lon", "lat"),crs = 4326, agr = "constant")
```

Select the columns that are of interest

```{r}
Data_4info<-Data_3spatial[,c("IDs","trip_number","dia_hora")]
```


# Map

Install the package **tmap**.

This packages allows you to create interactive maps. 

```{r}
library(tmap)
```

For making the map interactive, the mode **view** should be declared. 

```{r}
tmap_mode("view") 
```

To create your map with points you can use the arguments **tm_shape** and **tm_dots**

```{r}
tm_shape(Data_4info)+
   tm_dots()
```


To rename the information that shows, you can add the argument **popup.vars**

```{r}
tm_shape(Data_4info)+
   tm_dots(id = "IDs",col="yellow",popup.vars=c("trip_number"="trip_number","dt"="dia_hora"))
```


To separate between individuals you can also include them as separated layers. 

```{r}
tm_shape(subset(Data_4info,Data_4info$ID=='GPS01'))+
   tm_dots(id = "IDs",col="yellow",popup.vars=c("trip_number"="trip_number","dt"="dia_hora"))+
  tm_shape(subset(Data_4info,Data_4info$ID=='GPS03'))+
   tm_dots(id = "IDs",col="red",popup.vars=c("trip_number"="trip_number","dt"="dia_hora"))
```


Interact with you map

```{r, echo=FALSE}
library(fontawesome)
```

Using `r fa("plus", fill = 'black')` you can **zoom-in** to see the points.  
Using `r fa("minus", fill = 'black')` you can **zoom-out** to see the points.  

Using `r fa("layer-group", fill = 'black')` you can select which **layers** to show.  

Put the cursor on top of the dot `r fa("mouse-pointer", fill = 'black')`, when `r fa("hand-point-up", fill = 'black')` appears click on it and you should be able to see the ID, trip number and dt and time of the point. 

```{r}
tm_shape(subset(Data_4info,Data_4info$ID=='GPS01'))+
   tm_dots(id = "IDs",col="yellow",popup.vars=c("trip_number"="trip_number","dt"="dia_hora"))+
  tm_shape(subset(Data_4info,Data_4info$ID=='GPS03'))+
   tm_dots(id = "IDs",col="red",popup.vars=c("trip_number"="trip_number","dt"="dia_hora"))
```

Using tmap might save you the step to export your dataframe, and the import it into another program, particularly if you want to check a particular point and particular information associated to that point.


# Further reading

- [tmap](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)  
- [examples of use](https://r-tmap.github.io/tmap/index.html)  
- [book](https://r-tmap.github.io/tmap-book/index.html)  
- [geocomputation with R](https://r.geocompx.org/adv-map)
