---
title: "Kernel UD considerations"
description: |
  Some things to consider before making kernel density estimations.
author:
  - name: Miriam Lerma
    url: {}
date: 09-14-2022
preview: blog15.jpg
categories:
  - R
  - Biologging
  - English
  - Y2022
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    self_contained: false
---


# Intro

This post is to exemplify some considerations when calculating kernel density analyses. 

Before start calculating kernel density analyses, its useful to consider some sources of error that might change your results.

For the exercises, test data is from masked boobies. <br>
To access the data you have to install the package **sula**: devtools::install_github("MiriamLL/sula")

```{r}
#devtools::install_github("MiriamLL/sula")
library(sula)
GPS_raw<-(GPS_raw)
```

To manipulate the data we will use functions from the package **tidyverse**

```{r}
library(tidyverse)
```

For spatial manipulations we will use functions from the packages **sp** and **sf**

```{r}
library(sp)
library(sf)
```

For creating the polygons of kernel density we will use the package **adehabitathr**

```{r}
library(adehabitatHR)
```

## 1. Individuals


Some individuals might drive kernel density calculations in one or other direction as effect of different number of recordings (or days recorded) per individual.

### Unpaired

To illustrate this, lets see the calculations using together one individual sampled for 1 day and other for 5 days

#### Individual one day

```{r, echo=FALSE, eval=FALSE}
GPS_raw %>%
  filter(IDs=='GPS01') %>%
  distinct(DateGMT)
```

```{r}
ID_1day<-GPS_raw %>%
  filter(IDs=='GPS01') %>%
  filter(DateGMT %in% c('02/11/2017'))
```

#### Individual 5 days

```{r, echo=FALSE, eval=FALSE}
GPS_raw %>%
  filter(IDs=='GPS03') %>%
  distinct(DateGMT)
```

```{r}
ID_5days<-GPS_raw %>%
  filter(IDs=='GPS03')
```

#### Unpaired days

```{r}
Unpaired<-rbind(ID_1day,ID_5days)
```

Transform to spatial object.

```{r}
Unpaired<-as.data.frame(Unpaired)
coordinates(Unpaired) <- c("Longitude", "Latitude")
class(Unpaired)
```

Calculate kernelUD.

```{r}
UnpairedUD<-kernelUD(Unpaired[,3],h='href') 
```

Obtain polygons. 

```{r}
UnpairedUD95 <- getverticeshr(UnpairedUD, percent = 95, unout = c("m2"))
UnpairedUD50 <- getverticeshr(UnpairedUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Unpaired95<-st_as_sf(UnpairedUD95)
Unpaired50<-st_as_sf(UnpairedUD50)
```

```{r}
ggplot()+
  geom_sf(data = Unpaired95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Unpaired50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```

### Paired 

To compare, lets now see the kernel density calculated with these same individuals but recorded at similar number of days (3 days)

```{r}
ID_01<-GPS_raw %>%
  filter(IDs=='GPS01') %>%
  filter(DateGMT %in% c('02/11/2017','03/11/2017','04/11/2017','05/11/2017'))
```

```{r}
ID_03<-GPS_raw %>%
  filter(IDs=='GPS03') %>%
  filter(DateGMT %in% c('02/11/2017','03/11/2017','04/11/2017','05/11/2017'))
```

```{r}
Paired<-rbind(ID_01,ID_03)
```

Transform to spatial object.

```{r}
Paired<-as.data.frame(Paired)
coordinates(Paired) <- c("Longitude", "Latitude")
class(Paired)
```

Calculate kernelUD.

```{r}
PairedUD<-kernelUD(Paired[,3],h='href') 
```

Obtain polygons. 

```{r}
PairedUD95 <- getverticeshr(PairedUD, percent = 95, unout = c("m2"))
PairedUD50 <- getverticeshr(PairedUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Paired95<-st_as_sf(PairedUD95)
Paired50<-st_as_sf(PairedUD50)
```

```{r}
ggplot()+
  geom_sf(data = Paired95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Paired50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```

As you can see the resulting areas would be different.

To solve this problem, you might want to make sure to have tracking data of similar number of days or recordings.


## 2. Intervals

**Do you have similar recordings in time?**

If some devices have gaps, or record at different intervals, you might underestimate or overestimate specific areas.

For this example, lets see one individuals

```{r}
GPS01<-GPS_raw %>%
  filter(IDs=='GPS01')
```

### Gaps.

Using the column of hours, lets extract all the recordings after 5 pm.

```{r}
GPS01$Hour <- as.numeric(substr(GPS01$TimeGMT, 1, 2))
Gaps<-GPS01 %>%
  filter(Hour <= 17)
```

Transform to spatial object.

```{r}
Gaps<-as.data.frame(Gaps)
coordinates(Gaps) <- c("Longitude", "Latitude")
```

Calculate kernelUD.

```{r}
GapsUD<-kernelUD(Gaps[,3],h='href') 
```

Obtain polygons. 

```{r}
GapsUD95 <- getverticeshr(GapsUD, percent = 95, unout = c("m2"))
GapsUD50 <- getverticeshr(GapsUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Gaps95<-st_as_sf(GapsUD95)
Gaps50<-st_as_sf(GapsUD50)
```

```{r}
ggplot()+
  geom_sf(data = Gaps95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Gaps50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```

### Complete

In contrast, the kernel density calculations without gaps would give different results. 

```{r}
Complete<-GPS_raw %>%
  filter(IDs=='GPS01')
```

Transform to spatial object.

```{r}
Complete<-as.data.frame(Complete)
coordinates(Complete) <- c("Longitude", "Latitude")
```

Calculate kernelUD.

```{r}
CompleteUD<-kernelUD(Complete[,3],h='href') 
```

Obtain polygons. 

```{r}
CompleteUD95 <- getverticeshr(CompleteUD, percent = 95, unout = c("m2"))
CompleteUD50 <- getverticeshr(CompleteUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Complete95<-st_as_sf(CompleteUD95)
Complete50<-st_as_sf(CompleteUD50)
```

```{r}
ggplot()+
  geom_sf(data = Complete95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Complete50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```


To solve the problem with gaps, you can interpolate the data to fill the gaps and have [similar intervals](https://www.miriam-lerma.com/posts/2022-03-22-interpolation). 
However, caution should be taken if you have large gaps, it would create a line.

## 3. Behaviour

**Do you want to know the general areas that the animal used or just where it was feeding?**

It depends on your question, but if you are interested in specific behaviours, for example feeding areas, the kernel density analyses might be bring very different results than when using all movement data.

### Foraging

Here, we are using only areas where the animal was foraging. 

Load data

```{r}
GPS_raw<-as.data.frame(GPS_raw)
GPS01<-subset(GPS_raw,GPS_raw$IDs=='GPS01')
```

Use an specific period

```{r}
GPS_bc<-recortar_periodo(GPS_data=GPS01,
                                inicio='02/11/2017 18:10:00',
                                final='05/11/2017 14:10:00',
                                dia_col='DateGMT',
                                hora_col='TimeGMT',
                                formato="%d/%m/%Y %H:%M:%S")
```

Convert to the correct format

```{r}
GPS_bc$tStamp<-paste(GPS_bc$DateGMT,GPS_bc$TimeGMT)
GPS_bc$tStamp <- as.POSIXct(strptime(GPS_bc$tStamp,"%d/%m/%Y %H:%M:%S"),"GMT")
GPS_bc$lon<- as.numeric(GPS_bc$Longitude)
GPS_bc$lat<- as.numeric(GPS_bc$Latitude)
GPS_bc$id <- as.factor(GPS_bc$IDs)
```

Keep only the important columns

```{r}
GPS_bc<-GPS_bc %>%
  dplyr::select('id','tStamp','lon','lat')
```

Load the package

```{r, warning=FALSE}
library(EMbC)
```

Run the function 

```{r}
BC_clustering<-EMbC::stbc(GPS_bc[2:4],info=-1) 
```

Add the behavioral classifications

```{r}
GPS_bc$Behaviours<-(BC_clustering@A)
```

Rename it so you can understand what each behaviour means

```{r}
GPS_bc<-mutate(GPS_bc, BC = ifelse(GPS_bc$Behaviours == "1", "Resting",
                                        ifelse(GPS_bc$Behaviours == "2", "Intense foraging",
                                               ifelse(GPS_bc$Behaviours == "3", 'Travelling',
                                                      ifelse(GPS_bc$Behaviours == "4", "Relocating", 
                                                             "Unknown")))))


```

Filter to keep only foraging

```{r}
Foraging<-GPS_bc %>%
  filter(BC=='Intense foraging')
```

Transform to spatial object

```{r}
Foraging<-as.data.frame(Foraging)
coordinates(Foraging) <- c("lon", "lat")
```

Calculate kernelUD.

**Note** Here the href is of 0.0048 which is giving the error of *subscript out of bounds*
Lets then better calculate using other h value

```{r, eval=FALSE}
#ForagingUD<-kernelUD(Foraging[,3],h='href') 
#ForagingUD95 <- getverticeshr(ForagingUD, percent = 95, unout = c("m2"))
```

The new h value is of 0.01

```{r}
ForagingUD<-kernelUD(Foraging[,3],h=0.009) 
ForagingUD
```

Obtain polygons. 

```{r}
ForagingUD95 <- getverticeshr(ForagingUD, percent = 95, unout = c("m2"))
ForagingUD50 <- getverticeshr(ForagingUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Foraging95<-st_as_sf(ForagingUD95)
Foraging50<-st_as_sf(ForagingUD50)
```

```{r}
ggplot()+
  geom_sf(data = Foraging95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Foraging50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```

### All recordings
Here, we are using all the areas. 

```{r}
GPS_bc<-GPS_bc %>%
  dplyr::select('id','tStamp','lon','lat')
```

```{r}
Behas<-as.data.frame(GPS_bc)
coordinates(Behas) <- c("lon", "lat")
```

Calculate kernelUD.

```{r, message=FALSE, warning=FALSE}
BehasUD<-kernelUD(Behas,h='href') 
BehasUD
```

Obtain polygons. 

```{r}
BehasUD95 <- getverticeshr(BehasUD, percent = 95, unout = c("m2"))
BehasUD50 <- getverticeshr(BehasUD, percent = 50, unout = c("m2"))
```

Here you can check on you polygons visually.

```{r}
Behas95<-st_as_sf(BehasUD95)
Behas50<-st_as_sf(BehasUD50)
```

```{r}
ggplot()+
  geom_sf(data = Behas95,color='#006d77',fill = "#006d77",alpha=0.3,size=1)+
  geom_sf(data = Behas50,color='#5f0f40',fill = "#5f0f40",alpha=0.3,size=1)+
  labs(x = "Longitude", y="Latitude")+
  theme_bw()
```
If you want to classify the behaviour, please check the post on [EmBC](https://www.miriam-lerma.com/posts/2021-11-25-embc/)





