---
title: "Time overlaps"
description: |
  Find overlapping time periods
author:
  - name: Miriam Lerma
    url: {}
date: 02-4-2022
preview: blog7.jpg
categories:
  - R
  - tracking
  - timestampt
  - English
output:
  distill::distill_article:
    self_contained: false
---

# Intro

This post is about how to find overlapping times between your tracked individuals. <br>
This might help you to: <br>
- Identify periods where you have **gaps** for each individual, or <br>
- To **select which period is comparable among individuals** that might have been tracked on different periods <br>


# Data
To replicate the exercise, load data from the package 'sula'. <br>
For accessing the data, you need to have the package installed. 

```{r}
#devtools::install_github("MiriamLL/sula")
library(sula)
```

The data is from 10 tracked individuals.

```{r}
GPS_raw<-(GPS_raw)
```

# Plot

For this exercise you should install or have installed the package ggplot, and do not forget to 'call it'.

```{r}
library(ggplot2)
```


In the example data, time and date are in separated columns, therefore the first step is to join those columns.

```{r}
GPS_raw$DateTime<-paste(GPS_raw$DateGMT,GPS_raw$TimeGMT)
```

For plotting times, it is important that the column with the timestamps is in a time format.
Be cautious with the format of your time stamps.

```{r}
GPS_raw$DateTime<-as.POSIXct(strptime(GPS_raw$DateTime, "%d/%m/%Y %H:%M:%S"),tz = "GMT")
```


Now for plotting, you use the timestamp for the x axis and the IDs of each individual as the y axis (I consider that it is easy to see this way).
Also if you want each individual to be plotted in different color then you also add the name of the column after the argument color=.

```{r}
Plot_tracking<-ggplot(data=GPS_raw,
                      aes(x=DateTime,
                          y=IDs,
                          color=IDs))+
    geom_point(size=3)+
    geom_line()+
    theme_bw()
Plot_tracking
```

Because is a ggplot, many things can be further customize.

## Modify legend
Such as the *legends*

```{r}
Plot_tracking<-Plot_tracking+
    theme(legend.position='none')+
    labs(x = "", y = "Individuals") +
        theme(axis.text.x=element_text(angle=60, hjust=1))+
    theme(plot.title = element_text(size = 10, face = "bold"))
Plot_tracking
```

## Change breaks
You can also select how often you want the breaks, or how detailed the x axis legend should be.

For one day breaks:

```{r}
Plot_tracking + scale_x_datetime(date_labels = "%d %b %Y",date_breaks = "1 day")
```

For one week breaks:

```{r}
Plot_tracking + scale_x_datetime(date_labels = "%d %b %Y",date_breaks = "1 week")
```

For one month breaks:

```{r}
Plot_tracking + scale_x_datetime(date_labels = "%d %b %Y",date_breaks = "1 month")
```


## Add lines
You can also mark separations between periods using a line.

To do so, first you have to make the period where you want to make the line as POSIXct

```{r}
PA<-as.POSIXct(strptime("2017-11-07 00:00:00", "%Y-%m-%d %H:%M:%S"),tz = "GMT")
```

Then you add the line to the ggplot using the function geom_vline (v from vertical line)
Use the period you have interest on getting the line as xintercept, and
Linetypes option can be seen [here](http://www.cookbook-r.com/Graphs/Shapes_and_line_types/).
The color depends on your preference, here I use red to make it easy to see at first glance.

```{r}
Plot_tracking+
  geom_vline(xintercept=PA,
                linetype=4, colour="red")
```

## Add rectangle

Now if you want to add a rectangle, there are many option but I prefer to use annotation

First I create an object with the timestamps of where I want the rectangle to begging and end.

```{r}
PA<-as.POSIXct(strptime("2017-11-01 00:00:00", "%Y-%m-%d %H:%M:%S"),tz = "GMT")
PB<-as.POSIXct(strptime("2017-11-07 00:00:00", "%Y-%m-%d %H:%M:%S"),tz = "GMT")
```

Then I add it to the plot.
The xmin is the time where the rectangle will start, and the xmax where it would end.
Because I want the rectangle to cover the area I used -Inf and Inf
The color is in fill, and the alpha is to have some transparency (the lower the value the more transparent will be, and the values go from 0 to 1).

```{r}
Plot_tracking + annotate("rect",
                         xmin = PA, xmax = PB, 
                         ymin = -Inf, ymax = Inf,  
                         fill = "blue", alpha=0.1)
```
# Export

To export your plot, you can use the argument *ggsave*

The simplest is just to give it a name to your plot, and ggsave will save the latest plot you have created, you just have to give it a name and an extention.

Some options are: png, eps, pdf, jpeg, tiff, bmp or svg

```{r, eval=FALSE}
ggsave(filename = "~/Figure.png")
```

Note that the figure would be saved in the same folder where you have the Rmd. If you want an specific folder, change it in the file name.

```{r, eval=FALSE}
ggsave(filename = "Results/Figure.png")
```

If you want to create several figures and want to export just specific ones, you can also use the name of the plot to export only that one.

```{r, eval=FALSE}
ggsave(Plot_tracking, 
       filename = "~/Figura.png")
```

Also, if you want specific widths and heights, or even dpi, you can also include it. 
Just note that the size of the letters might need to be adjusted.

```{r, eval=FALSE}
ggsave(Plot_tracking, 
       filename = "~/Figura.png",
       width = 24, 
       height = 24, 
       units = "in",
       dpi = 500)
```

# Recommended sources

- [ggplot2](https://ggplot2.tidyverse.org/index.html)
- [ggsave](https://ggplot2.tidyverse.org/reference/ggsave.html) 
