---
title: "Identify events"
description: |
  Assign a number to each event.
author:
  - name: Miriam Lerma
    url: {}
date: 2024-09-23
preview: blog37_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2024
---

# Intro

To identify between different events assigning an ID is essential.  
This post shows an option to assign a trip number to locations of animals. 

# Data 

For the exercises, test data is from Masked boobies. <br>
To access the data the package **sula** needs to be installed: devtools::install_github("MiriamLL/sula")

```{r}
#devtools::install_github("MiriamLL/sula")
library(sula)
```

This data frame contains data from 10 individuals.

```{r}
unique(sula::GPS_raw$IDs)
```
Using functions from the package **tidyverse**, one of the individuals can be selected. 

```{r}
library(tidyverse)
```

```{r}
GPS_one<-GPS_raw %>% filter(IDs=='GPS08')
```

The original data can be checked using functions from the package **ggplot2**.

```{r}
library(ggplot2)
```

All locations are presented, but events can not be identify yet. 

```{r}
Plot_original<-ggplot(GPS_one, aes(x=Longitude, y=Latitude)) + 
  geom_point()+
  scale_x_continuous(labels = function(x) paste0(-x, '\u00B0')) +
  scale_y_continuous(labels = function(x) paste0(-x, '\u00B0'))  +
  ggtitle('Tracking data')+xlab('Longitude')+ylab('Latitude')+
  theme_bw()+
  theme(
    legend.position='none',
    panel.background = element_rect(fill = '#edf2f4'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1.5))
Plot_original
```


# Remove central locations

Here, an event would be defined as each time the animal leaves its central location.  

Provide information of the central location

```{r}
Central_location<-data.frame(Longitude=-109.4531, Latitude=-27.20097)
```

Transform the locations into spatial data.

```{r}
GPS_spatial <- GPS_one
sp::coordinates(GPS_spatial) <- ~Longitude + Latitude
sp::proj4string(GPS_spatial) = sp::CRS("+init=epsg:4326")
GPS_spatial<-sf::st_as_sf(GPS_spatial)
```

The buffer function is explained in the previous [post](https://www.miriam-lerma.com/posts/2024-08-01-removelocs/).  
The function uses the central location to create a buffer.  
A number "1" shows that the location is inside the buffer and an NA that is outside the buffer.   
The information is added on a column in the data frame.  

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  central_spatial<-sf::st_as_sf(central_spatial)
  buffer_dist<-buffer_km*1000
  central_buffer<-sf::st_buffer(central_spatial, buffer_dist)
  return(central_buffer)
  }
This_buffer<-create_buffer(central_point=Central_location,buffer_km=1)
GPS_over<-sapply(sf::st_intersects(GPS_spatial,This_buffer), function(z) if (length(z)==0) NA_integer_ else z[1])
GPS_one$trip <- as.numeric(GPS_over)
```

Replace the NAs (when the animal is outside the buffer) for a 0. This is not strictly necessary but might facilitate the interpretation of the column.

```{r}
GPS_one$trip[is.na(GPS_one$trip)] <- 0
```

Alternatively, the numbers can be replaced by characters.

```{r}    
GPS_one$trip <- gsub("1", "At_central_locations", GPS_one$trip)
GPS_one$trip <- gsub("0", "At_trip", GPS_one$trip)
```

# Assign trip number

Add a new column with a sequential number.

```{r}
GPS_one$sequential_number<-as.integer(paste(seq(1:as.numeric(nrow(GPS_one)))))
```

Remove the locations inside the central locations using the function **subset**

```{r}
GPS_trips<-subset(GPS_one, GPS_one$trip == "At_trip")
```

Add a number of event everytime that the sequence is broken. 

```{r}
GPS_trips$trip_number <- (cumsum(c(1L, diff(GPS_trips$sequential_number)) != 1L))+ 1
```

Add more zeros using the function **str_pad** if there are more than ten trips. 

```{r}
GPS_trips$trip_number <- stringr::str_pad(GPS_trips$trip_number ,  2, pad = "0")
```

Add text for easier interpretation. 

```{r}
GPS_trips$trip_number<-paste0("Trip_", GPS_trips$trip_number )
```

Check the number of events (or trips) using the function **unique**.

```{r}
unique(GPS_trips$trip_number)
```

# Plot

By adding the argument **color=trip_number**, the different events can now be easily observed. 

```{r}
Plot_trips<-ggplot(GPS_trips, aes(x=Longitude, y=Latitude, color=trip_number)) + 
  geom_point()+
  geom_point(data=Central_location,aes(x = Longitude,y= Latitude), 
             color="black", fill="black",shape=16,size=5,stroke=1.5)+
  theme_bw()+
  scale_x_continuous(limits=c(-109.7,-108.9),labels = function(x) paste0(-x, '\u00B0')) +
  scale_y_continuous(labels = function(x) paste0(-x, '\u00B0'))  +
  ggtitle('Trip number')+xlab('Longitude')+ylab('Latitude')+labs(color='')+
  theme(
    panel.background = element_rect(fill = '#edf2f4'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=1.5),
    legend.background = element_rect(colour = "transparent", fill = "transparent"),
    legend.position = c(0.12,0.75))
Plot_trips
```

```{r, echo=FALSE, eval=FALSE}
library(here)
library(patchwork)
here()
ggsave(Plot_original+Plot_trips,
       filename = paste0(here(),"/_posts/2024-09-19-identifyevents/blog37.jpg"),
       width = 9, 
       height = 4, 
       units = "in",
       dpi = 300)
```
# Further reading

Functions **count_trips** from the package sula can be used to run loops. 



