---
title: "ezzs"
description: |
  A short description of the post.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-03-12
output:
  distill::distill_article:
    self_contained: false
---


# EZZ and protected areas

# Packages

```{r}
library(here)
library(ggplot2)
library(ggspatial)
library(plotly)
library(tidyverse)
```


```{r}
#devtools::install_github("MiriamLL/GermanNorthSea")
library(GermanNorthSea)
```

# Data

To download the original data go to [GeoSeaPortal](https://www.bsh.de/DE/THEMEN/Offshore/Meeresraumplanung/Raumordnungsplan_2021/raumordnungsplan-2021_node.html)

The zip contains several shapefiles including NatureConservation

Other option for Natura 2000 shapefiles is [European Environment Agency](https://www.eea.europa.eu/en/datahub/datahubitem-view/6fc8ad2d-195d-40f4-bdec-576e7d1268e4)

This file contains a more complete data from the natura 2000 sites.

For using the SCA (Special Conservation Areas) in Germany.

# Plot

```{r, fig.height=8, fig.width=8}
Germany_plot<-ggplot(WindFarms) + # Gives CRS
  
  # Fill colors
  geom_sf(data = German_EEZ, colour = "black", fill= '#56C1D5', lwd = 0.5)+
  
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#84a98c', alpha=0.5,lwd = 0.5)+
  geom_sf(data = German_SCA, colour = "#3d6d22", fill= '#84a98c', alpha=0.5,lwd = 0.5)+
 
  # Line colors
  geom_sf(data = German_EEZ, colour = "black", fill= NA, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "black", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = '#9a8c98', fill = '#9a8c98')+
  
  # Line colors
  geom_sf(data = German_coast, colour = "black", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#84a98c', alpha=0.5,lwd = 0.5)+
  geom_sf(data = German_coast, colour = "black", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  
  # Annotations
  annotation_north_arrow(location ="bl", which_north = "true", 
        pad_x = unit(0.05, "in"), 
        pad_y = unit(0.05, "in"),
        
        style = north_arrow_fancy_orienteering)+
  
  
  theme(
  legend.position='none',
  legend.spacing.y = unit(0.05, 'cm'),
  legend.text=element_text(size=10),
  legend.background = element_rect(fill='transparent',colour ="transparent"),
  legend.box.background = element_rect(fill='transparent',colour ="transparent"),
  legend.key = element_rect(fill = "transparent", colour = "transparent"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.background = element_rect(fill = '#AFDEE8'))+
  
  xlab('Longitude')+ylab('Latitude')+
  
  theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3900000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(left = "N", bottom = 'E'))

Germany_plot
```

```{r, fig.height=6, fig.width=8}
Germant_labels<-Germany_plot+
  #Dogger bank
  annotate("text", x = 3900000+50000+50000+30000-15000, y = 3479000+150000, label = "Dogger\n Bank",colour = '#3d6d22',size=4.5)+
  annotate("segment", x = 4030000-60000, xend = 4030000-45000+11000, 
           y = 3479000+150000, yend = 3479000+150000,
           colour = '#3d6d22', alpha=0.5)+
  
  ## Sylt
  annotate("text", x = 3900000+50000+50000+30000+129000, 
           y = 3479000+120000-1000+18000+15000, label = "Sylt \n Outer \n Reef",colour = '#3d6d22',size=4.5)+
  annotate("segment", x = 4159000, xend = 4159000, 
           y = 3618000-8000-8000-13000+18000, yend = 3566000-1800+1000,
           colour = '#3d6d22', alpha=0.5)+
  
  # Borkum
  annotate("text", x = 4159000-180000-18000, y = 3479000-38000-10000-10000, 
           label = "Borkum \n Reef \n Ground",colour = '#3d6d22',size=4.5)+
  annotate("segment", x = 4159000-110000-70000, xend = 4159000-110000+22000+6000, 
           y = 3479000-38000-10000-10000,yend = 3479000-38000-10000-10000,
           colour = '#3d6d22', alpha=0.5)
Germant_labels
```


```{r, echo=FALSE, eval=FALSE}
## Export
ggsave(Germant_labels,
       filename = paste0(this_folder,"/chl_GermanNorthSea.jpg"),
       width = 8, 
       height = 8, 
       units = "in",
       dpi = 300)
```

# References

[EEZ]()
[Natura]()