---
title: "Light pollution"
description: |
  Create a map including light pollution.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-02-10
preview: blog42_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---

# Data

Download data from [Light pollution map](https://www.lightpollutionmap.info/#zoom=3.32&lat=48.8875&lon=2.7774&state=eyJiYXNlbWFwIjoiTGF5ZXJCaW5nUm9hZCIsIm92ZXJsYXkiOiJ2aWlyc18yMDE3Iiwib3ZlcmxheWNvbG9yIjpmYWxzZSwib3ZlcmxheW9wYWNpdHkiOjYwLCJmZWF0dXJlc29wYWNpdHkiOjg1fQ==)
- By selecting the tools, then the square.  
- Create square in the area of interest and wait until information appears, then click in the download icon.  

To read from file.

Select the directory where the information is.

```{r, eval=FALSE}
library(here)
here()
your_folder<-paste0(here(),"/_posts/2024-10-18-lightpollution")
list.files(your_folder)
```

```{r, eval=FALSE}
your_file<-paste0(your_folder,"/viirs_npp_202300b.tif")
```

Use the package **terra** to use the function **rast**.  

```{r, eval=FALSE}
library(terra)
```

```{r, eval=FALSE}
Light_pollution<-rast(your_file)
```

Then convert to data frame. 

```{r, eval=FALSE}
Light_pollution_df <- as.data.frame(Light_pollution, xy = TRUE)
beepr::beep(sound=1)
```

```{r}
library(tidyverse)
```

```{r}
Light_pollution_df<-Light_pollution_df %>%
  rename(radiance=3)
```

```{r}
range(Light_pollution_df$radiance)
```

```{r}
Light_pollution_df<-Light_pollution_df %>%
  mutate(radiance=radiance/1000)
```

```{r}
range(Light_pollution_df$radiance)
```

# Plot

## Basic plot

Use **ggplot** to create your plot. 

```{r}
library(ggplot2)
```

```{r, fig.height=6, fig.width=6}
Plot_light<-ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  geom_sf(data = German_land, colour = 'black', fill = '#335c67', alpha=0.05)+
  scale_fill_distiller(palette = "Spectral")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_light+
  guides(fill=guide_legend(title="Radiance"))
```

## Adjust fill

```{r}
range(Light_pollution_df$radiance)
```

```{r}
hist(Light_pollution_df$radiance)
```

Include **scale_fill_gradientn** note the n for continuous variables

```{r, eval=FALSE, echo=FALSE}
color_A <-"#000000"
color_B <-"#051637"
color_C <-"#20998F"
color_D <-"#10BA47"
color_E <-"#55E700"
color_F <-"#AAF300"
color_G <-"#FFFF00"
color_H <-"#F9DF06"
color_I <-"#F3C00C"
color_J <-"#C2523C"
color_K <-"#B92927"
color_L <-"#B00012"
your_palette<-c(color_A,color_B,color_C,color_D,color_E,color_F,color_G,color_H,color_I,color_J,color_K,color_L)
```


```{r, fig.height=6, fig.width=6, eval=FALSE, echo=FALSE}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  scale_fill_gradientn(
    colours = your_palette,
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)))
```

## Add breaks

```{r, fig.height=6, fig.width=6, eval=FALSE, echo=FALSE}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  scale_fill_gradientn(
    colours = your_palette,
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)),
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.2)
    )
```

## Legend 

```{r, fig.height=6, fig.width=6, eval=FALSE, echo=FALSE}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  scale_fill_gradientn(
    colours = your_palette,
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)),
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.2)
    )+
  theme(legend.position='bottom',
        legend.key.size = unit(1.8, "cm"))
```


## Guide

Add guide= legend

Now the values are discrete values

```{r, eval=FALSE, echo=FALSE}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  scale_fill_gradientn(
    colours = your_palette,
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)),
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.2),
    guide="legend"
    )+
  theme(legend.position='bottom')+ guides(fill = guide_legend(nrow = 1))
```



## Add land

Read package German North Sea for extra shapefiles

```{r}
library(GermanNorthSea)
```

Use the package sf for adjusting the CRSs

```{r}
library(sf)
```

Make sure is in the same CRS. 

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```

```{r, fig.height=6, fig.width=6}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  geom_sf(data = German_land, colour = 'black', fill = '#e5e5e5', alpha=0.05)+
  scale_fill_gradientn(
    colours = your_palette,
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)),
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.2)
    )+
  coord_sf(xlim = c(4,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


## Other colors

```{r, fig.height=6, fig.width=6}
ggplot() +
  geom_raster(data = Light_pollution_df, aes(x = x, y = y, fill = radiance)) +
  geom_sf(data = German_land, colour = 'black', fill = '#e5e5e5', alpha=0.08)+
  scale_fill_gradientn(
    name= "Radiance",
    colours = c("#000000","#051637","#20998F","#10BA47","#AAF300","#FFFF00","#F9DF06","#F3C00C","#C2523C","#B92927","#B00012"),
    limits  = c(0, 2.2),
    values  = scales::rescale(c(0,0.10,0.25,0.30,0.35,1.50,2.00,3.00,6.00,15.0,30.0,40.0,65.0,80.0,150)),
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0,1.2)
    )+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(4,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


# Further read

[Viirs](https://www.sciencedirect.com/topics/earth-and-planetary-sciences/viirs)
[Virrs Mission](https://ladsweb.modaps.eosdis.nasa.gov/missions-and-measurements/viirs/)

Mostly at land
[Dark Sky](https://darksky.org/news/eyes-in-the-sky-exploring-global-light-pollution-with-satellite-maps/)
[Blue MArble](https://blue-marble.de/nightlights/)

At Sea
[Global atlas of artifical light](https://online.ucpress.edu/elementa/article/9/1/00049/119144/A-global-atlas-of-artificial-light-at-night-under)


