---
title: "Plot distance to coast in R"
description: |
  Create a map with distance to coast information in R.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-02-20
preview: blog42_portadas.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---

# Intro

Distance to coast gives us information on the distance (in meters) from one point at sea to the nearest coast.

This post includes:  
- Download raster data   
- Read and subset raster data  
- Plot raster data  

# Distance to coast

To download:   
- Access [OceanColor NASA](https://oceancolor.gsfc.nasa.gov/resources/docs/distfromcoast/)  
- Select **the interpolated 0.01-degree GeoTiff packed together with a brief description file.**  
- Unzip information.  

# Read and subset

To read from file.

Select the directory where the information is.

```{r, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/data")
this_file<-paste0(this_folder,"/GMT_intermediate_coast_distance_01d.tif")
```

Use the package **terra** to use the function **rast**.  
Then convert to data frame.  

```{r, eval=FALSE}
library(terra)
DistCoast_file<-rast(this_file)
DistCoast_dataframe <- as.data.frame(DistCoast_file, xy = TRUE)
beepr::beep(sound=1)
```

The file is quite large, so I recommend to subset the data to the area of interest.  

Here I select the area close to the German North Sea. 

```{r, eval=FALSE}
library(tidyverse)
DistCoast_dataframe_sub <-DistCoast_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)%>%
  rename(Dist=3) %>%
  mutate(Dist = as.numeric(Dist))
beepr::beep(sound=1)
```

To save as rda would make the reading a lot faster.

```{r,eval=FALSE}
German_distancecoast<-DistCoast_dataframe_sub
save(German_distancecoast, file="German_distancecoast.rda")
```

From the package.  

Alternatively, access the information from the package **GermanNorthSea**.

```{r}
library(GermanNorthSea)
```

```{r}
DistCoast_dataframe<-GermanNorthSea::German_distancecoast
```

# Plot

To plot adding land.

```{r}
library(sf)
```

Make sure is in the same CRS. 

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```

To exclude information on land. 

```{r}
library(tidyverse)
```

```{r}
DistCoast_dataframe_sub<-DistCoast_dataframe %>%
  filter(Dist > -20)
```

Use **ggplot** to create your plot. 

```{r, fig.width=4, fig.height=4}
Plot_distance<-ggplot() +
  geom_raster(data = DistCoast_dataframe_sub, aes(x = x, y = y, fill = Dist)) +
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "rocket")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_distance+
  guides(fill=guide_legend(title="Distance to coast"))
```


```{r, echo=FALSE, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2024-10-15-distancetocoast")
ggsave(Plot_distance,
       filename = paste0(this_folder,"/blog42.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

- [Package GermanNorthSea](https://github.com/MiriamLL/GermanNorthSea)
- [Package terra](https://rspatial.github.io/terra/) 
- [Package sf](https://r-spatial.github.io/sf/)
- [OceanColor NASA](https://oceancolor.gsfc.nasa.gov/resources/docs/distfromcoast/) 

