---
title: "Fishing effort in the German North Sea"
description: |
  Create a map with sst information from Global fishing watch in R. 
author:
  - name: Miriam Lerma
    url: {}
date: 2025-03-03
preview: Y2025_apr_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---


# Intro

[Global fishing watch](https://globalfishingwatch.org/datasets-and-code/)

The Global Fishing Watch map is the first open-access online tool for visualization and analysis of vessel-based human activity at sea. Anyone with an internet connection can access the map to monitor global fishing activity from 2012 to the present for more than 65,000 commercial fishing vessels that are responsible for a significant part of global seafood catch.

We use data that is broadcast using the automatic identification system (AIS) and collected via satellites and terrestrial receivers. We then combine this information with vessel monitoring system data provided by our partner countries. We apply our fishing detection algorithm to determine “apparent fishing effort” based on changes in vessel speed and direction. The heat map grid cell colors show how much fishing happened in that area, allowing for precise comparison.


Login using your email.

[Fishing effort](https://globalfishingwatch.org/map?access-token=wi1_Ffdm-AjT1k_urFVmOPTMqMXrKQ)


Tif is missing gear type data
Csv is missing lat and long

# Data

I want data of apparent fishing effort (AIC)
To use the data, you need to register (is free of use).
My first step is to limit the data to April and May of 2018.
[Video](https://www.youtube.com/watch?v=iWkgapKbJWg)

# Raster

```{r}
library(here)
here()
your_folder<-paste0(here(),"/_posts/2025-03-03-fishingdata")
list.files(your_folder)
```

```{r}
your_file<-paste0(your_folder,"/public-global-fishing-effort-v3.0.tif")
```

Use the package **terra** to use the function **rast**.  

```{r}
library(terra)
```

```{r}
Fishing_tif<-rast(your_file)
```

```{r}
class(Fishing_tif)
```

### Plot tif data

Install package tidyterra

```{r}
library(tidyterra)
```
```{r}
library(ggplot2)
```

```{r}
ggplot() +
  geom_spatraster(data = Fishing_tif) +
  facet_wrap(~lyr, ncol = 2)+
  theme_bw()
```

### Tif To data frame

Then convert to data frame. 

```{r}
Fishing_df<- as.data.frame(Fishing_tif, xy = TRUE)
beepr::beep(sound=1)
```

```{r}
head(Fishing_df)
```


### Subset 

```{r}
library(tidyverse)
```

Use **filter** to subset your data 

```{r}
Fishing_sub <-Fishing_df  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)%>%
  rename(FishingEffort=3) 
```

```{r}
mean(Fishing_sub$FishingEffort)
```

## Plot

```{r}
library(sf)
```

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
German_EEZ<-st_transform(GermanNorthSea::German_EEZ, 4326)
```

Use **ggplot** to create your plot. 

```{r, fig.width=4, fig.height=4}
Fishing_plot<-ggplot() +
  geom_tile(data = Fishing_sub, aes(x = x, y = y, fill = FishingEffort)) +
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_EEZ, colour = 'black', fill = 'transparent')+
  theme_bw()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Fishing_plot
```

# Csv

```{r}
list.files(paste0(your_folder,"/layer-activity-data-0"))
```

```{r}
csv_file<-paste0(your_folder,"/layer-activity-data-0","/public-global-fishing-effort-v3.0.csv")
```

```{r}
Fishing_csv<-read_csv(csv_file)
```


```{r}
Fishing_csv %>%
  distinct_all()
```


```{r}
Fishing_csv %>%
  group_by(Geartype)%>%
  tally()
```


# Export

```{r, echo=FALSE, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2025-03-03-fishingdata")
ggsave(Fishing_plot,
       filename = paste0(this_folder,"/blog54.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```


# Further read

[Working with public data](https://globalfishingwatch.org/data/downloadable-public-data-in-r/#)
[GitHub with outdate instructions](https://github.com/GlobalFishingWatch/paper-global-footprint-of-fisheries/blob/master/data_documentation/fishing_effort_byvessel.md)
[Tracking the global footprint of fisheries](https://www.science.org/doi/10.1126/science.aao5646)