---
title: "CHL German North Sea in R"
description: |
  Create a map with sst information from Copernicus in R.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-06-15
preview: blog56_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---



# Intro

Sea surface temperatures gives us information on Celsius around an area.

This post includes:  
- Download raster data   
- Read and subset raster data  
- Plot raster data  

# Data

Download raster data

Copernicus Marine Data Store

To download [visit webpage](https://data.marine.copernicus.eu/products)  
You need to register first  

Go to:  
- Click on the icon of download  
- [Click on download](https://help.marine.copernicus.eu/en/articles/4469993-how-to-download-copernicus-marine-products#h_ec9344c750)  
- Here I selected Mass concentration of chlorophyll a in sea water
- Add your coordinates here I use 1 to 10 and 50 to 60    
- Add you time scale here I use April and May 2018  
- I select the depths form 0 to 60 because I am using birds data  


Read raster data from a file

```{r, include=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2025-06-15-chlorophyll")
list.files(this_folder)
```

```{r, include=FALSE}
this_file<-paste0(this_folder,"/cmems_mod_arc_bgc_my_ecosmo_P1M_1729862405550.nc")
this_file
```

Load the [terra package](https://pjbartlein.github.io/REarthSysSci/netcdf_terra.html) for reading raster data

```{r}
library(terra)
```

The function **rast** helps to read raster data - replacing package raster

```{r}
CHL_file<-rast(this_file)
```

Change to data frame

```{r}
CHL_dataframe <- as.data.frame(CHL_file, xy = TRUE)
```

# Subset raster data

Load the package **tidyverse** 

```{r}
library(tidyverse)
```

Use **filter** to subset your data to your geographical area

```{r}
CHL_sub <-CHL_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)
```

## CHL info

Data was collected almost every meter

Therefore there are many columns with data per depth

```{r}
ncol(CHL_sub)
```

Keep columns with the depth area

```{r}
CHL_depth<-CHL_sub %>%
  select(starts_with("chl"))
```

The SST data is collected almost for every meter

```{r}
Depths<-colnames(CHL_depth)
head(Depths)
```

```{r}
CHL_depth_perloc<-CHL_depth %>%
  rowwise()%>%
  mutate(mean_CHL = mean(c_across(where(is.numeric)),na.rm=TRUE),
         min_CHL = min(c_across(where(is.numeric)),na.rm=TRUE),
         max_CHL = max(c_across(where(is.numeric))),na.rm=TRUE)%>%
  relocate(mean_CHL,min_CHL,max_CHL)
```

## Check

```{r}
first_column<-CHL_depth[1,]
number_cols<-ncol(first_column)
number_cols
```


```{r, warning=FALSE, eval=FALSE}
long_values<-first_column %>% 
  pivot_longer(cols = 1:number_rows, 
               names_to = "type",
               values_to = "value")
```

```{r, eval=FALSE}
arrange_values<-arrange(long_values, desc(value))
head(arrange_values)
```


## Mean values

```{r}
CHL_sub$CHL<-CHL_depth_perloc$mean_CHL
```

```{r}
mean(CHL_sub$CHL)
```

# Plot

Load the package **ggplot2** 

```{r}
library(ggplot2)
```

Plot your data using **geom_raster**

Select [palette](https://ggplot2-book.org/scales-colour)

```{r, fig.width=6, fig.height=4}
ggplot() +
  geom_raster(data = CHL_sub , aes(x = x, y = y, fill = CHL)) +
  scale_fill_distiller(palette = "YlOrBr")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```

Add land for reference

Load the package **sf**

```{r}
library(sf)
```

Use the function **st_transform** to convert to the same CRS

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```

Add the land to the plot using **geom_sf**

```{r, fig.width=6, fig.height=6}
Plot_chl<-ggplot() +
  geom_tile(data = CHL_sub , aes(x = x, y = y, fill = CHL)) +
  scale_fill_distiller(palette = "YlOrBr")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  theme_void()+
  theme(legend.position='left')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_chl+ ggtitle('Sea surface temperature from 0 to 60 m depth \n in the German North Sea in \n spring (April and May montly composites)')
```



```{r, echo=FALSE, eval=FALSE}
ggsave(Plot_chl,
       filename = paste0(this_folder,"/chl_GermanNorthSea.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

- [Package GermanNorthSea](https://github.com/MiriamLL/GermanNorthSea)
- [Package terra](https://rspatial.github.io/terra/) 
- [Package sf](https://r-spatial.github.io/sf/)

- [Examples](https://marine.copernicus.eu/services/user-learning-services/using-r-work-copernicus-marine-data)
- [How to download copernicus marine products](https://help.marine.copernicus.eu/en/articles/4469993-how-to-download-copernicus-marine-products#h_ec9344c750)


Comparing with other studies seems to fit

- [OSPAR](https://oap.ospar.org/en/ospar-assessments/intermediate-assessment-2017/pressures-human-activities/eutrophication/chlorophyll-concentrations/)
- [Copernicus](https://os.copernicus.org/articles/20/1247/2024/)

# End of document