---
title: "Remove undesired locations"
description: |
  Create a buffer to remove locations.
author:
  - name: Miriam Lerma
    url: {}
date: 2024-08-01
preview: blog36_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2024
---

# Intro

When studying animals using GPSs, we often need to remove their central location.
Here, I am sharing a function I created to eliminate all the locations within an area.

# Data 

For the exercises, test data is from masked boobies. <br>
To access the data you have to install the package **sula**: devtools::install_github("MiriamLL/sula")

```{r}
#devtools::install_github("MiriamLL/sula")
library(sula)
```

This data frame contains data from 10 individuals.

```{r}
unique(sula::GPS_raw$IDs)
```

To load it into the environment.

```{r}
GPS_ten<-GPS_raw
```

Plot your data

```{r}
library(tidyverse)
```

Here you can see all the recorded locations using the GPSs.

```{r, fig.width=4, fig.height=4}
Original_plot<-ggplot()+
  geom_point(data = GPS_ten,
            aes(x=Longitude,y = Latitude),color='red', size=0.5)+
  scale_x_continuous(labels = function(x) paste0(-x, '\u00B0')) +
  scale_y_continuous(labels = function(x) paste0(-x, '\u00B0'))  +
  xlab('Longitude')+ylab('Latitude')+
  theme(
    panel.background = element_rect(fill = '#edf2f4'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),legend.position='none',
    panel.border = element_rect(colour = "black", fill=NA, size=1.5)
  )
Original_plot
```


# Create a buffer

Select a location

```{r}
This_point<-data.frame(Longitude=-109.99,Latitude=-27.02)
```

Create a buffer

I created this function to create a buffer around a point

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  central_spatial<-sf::st_as_sf(central_spatial)
  buffer_dist<-buffer_km*1000
  central_buffer<-sf::st_buffer(central_spatial, buffer_dist)
  return(central_buffer)
  }
```

The parameters to give are the kilometers and the central point

```{r, warning=FALSE}
This_buffer<-create_buffer(central_point=This_point,buffer_km=20)
```

```{r}
class(This_buffer)
```


Here you can see the buffer you created using the point (or central location)

```{r, fig.width=4, fig.height=4}
Buffer_plot<-ggplot()+
  geom_point(data = GPS_ten,
            aes(x=Longitude,y = Latitude),color='red',size=0.5)+
  geom_point(data=This_point,
             aes(x=Longitude,y=Latitude),color='blue')+
  geom_sf(data=This_buffer,colour='blue', fill='transparent', linetype='dashed')+
  scale_x_continuous(labels = function(x) paste0(-x, '\u00B0')) +
  scale_y_continuous(labels = function(x) paste0(-x, '\u00B0'))  +
  xlab('Longitude')+ylab('Latitude')+
  theme(
    panel.background = element_rect(fill = '#edf2f4'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),legend.position='none',
    panel.border = element_rect(colour = "black", fill=NA, size=1.5)
  )
Buffer_plot
```



Using the function from the package **sp** you can create an spatial object using your GPS data

```{r}
GPS_sp <- GPS_ten
sp::coordinates(GPS_sp) <- ~Longitude + Latitude
sp::proj4string(GPS_sp) = sp::CRS("+init=epsg:4326")
GPS_sp<-sf::st_as_sf(GPS_sp)
```

```{r}
class(GPS_sp)
class(This_buffer)
```

Here the function over allows for identifying which location intersect with the buffer
To identify if the locations are in or out you can create a column

```{r}
GPS_over<-sapply(sf::st_intersects(GPS_sp,This_buffer), function(z) if (length(z)==0) NA_integer_ else z[1])
```

```{r}
GPS_ten$In_or_out <- as.numeric(GPS_over)
```

To remove the locations that were within the buffer you can use the function **filter** and **is.na** from the package **tidyverse**

```{r}
GPS_without <- GPS_ten %>%
  filter(is.na(In_or_out)==TRUE)
```

Here you can see that the locations inside the buffer were eliminated

```{r, fig.width=4, fig.height=4}
Filtered_plot<-ggplot()+
  geom_point(data = GPS_without,
            aes(x=Longitude,y = Latitude),color='red',size=0.5)+
  geom_point(data=This_point,
             aes(x=Longitude,y=Latitude),color='blue')+
  geom_sf(data=This_buffer,colour='blue', fill='transparent', linetype='dashed')+
  scale_x_continuous(labels = function(x) paste0(-x, '\u00B0')) +
  scale_y_continuous(labels = function(x) paste0(-x, '\u00B0'))  +
  xlab('Longitude')+ylab('Latitude')+
  theme(
    panel.background = element_rect(fill = '#edf2f4'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),legend.position='none',
    panel.border = element_rect(colour = "black", fill=NA, size=1.5)
  )
Filtered_plot
```


This function can be used to remove all the nest locations within a distance buffer.
You could also replace the polygon (here the buffer) for some land shapefiles.


# Compare

Using the package **patchwork** we can see the difference side by side.

```{r}
library(patchwork)
```

```{r, fig.width=10, fig.height=4}
Original_plot+Buffer_plot+Filtered_plot
```

```{r, echo=FALSE, eval=FALSE}
library(here)
here()
ggsave(Original_plot+Buffer_plot+Filtered_plot,
       filename = paste0(here(),"/_posts/2024-07-19-removelocs/blog36.jpg"),
       width = 10, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

Other functions that did the job:   
gBuffer from rgeos [deprecated](https://cran.r-project.org/web/packages/rgeos/index.html)  
Read more about [package migration](https://github.com/r-spatial/sf/wiki/migrating)



```{r, echo=FALSE, eval=FALSE}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  buffer_dist<-buffer_km/100
  central_buffer<-rgeos::gBuffer(central_spatial, width=1*buffer_dist)
  return(central_buffer)
}
GPS_ten$In_or_out <- as.numeric(sp::over(GPS_sp,This_buffer))
```



