---
title: "Day and night"
description: |
  Create a column with the classification day or night
author:
  - name: Miriam Lerma
    url: {}
date: 2022-12-04
preview: blog18.jpg
categories:
  - R
  - Tracking
  - English
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    self_contained: false
---


# Intro

This post is to create a column day or night.   


## Data
For the exercises, test data is from masked boobies. <br>
To access the data you have to install the package **sula**: devtools::install_github("MiriamLL/sula")


```{r}
#devtools::install_github("MiriamLL/sula")
library(sula)
GPS01<-(GPS_01)
```

## Transform time and date
```{r}
library(dplyr)
```

Identify column with time date data

```{r}
GPS01$dt <- as.POSIXct(strptime(GPS01$tStamp, "%Y-%m-%d %H:%M:%S"))
```


## Extract hours

Once is in time and date format, you can extract just the hours

```{r}
GPS01$hour <- as.numeric(substr(GPS01$dt, 12, 13))
```

## Classify day and night
Create a classification base on the time of sunset and sunrise.

```{r}
GPS01<-GPS01 %>%
 mutate(
    day_night = case_when(
      hour > 20 |  hour < 6 ~ "night",
      TRUE  ~ "day"
    ))
```

## Quantify proportions

```{r}
library(tidyverse)
```

Now you can quantify how much your animals were out at night

```{r}
GPS01 %>%
  group_by(day_night)%>%
  count()%>%
  pivot_wider(names_from = day_night, values_from = n)
```

I hope this might help you

## Further reads

Classify activities at night [in penguins](https://www.miriam-lerma.com/posts/bls7)

