---
title: "BLS7"
description: |
  Link to presentation.
preview: BLS_01.jpg
author:
  - name: Miriam Lerma
    url: {}
date: 08-03-2021
categories:
  - Seabird
  - presentation
  - English
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distilltools)
```


# Intro

This post is dedicated to my presentation for the Bio-Logging Symposium (https://www.bls7hawaii.com)

Diel foraging behaviour of Humboldt penguins *Spheniscus humboldti* at Tilgo Island, Northern Chile

Materials are underdevelopment 👩🏽‍🔧.

```{r icon-links1,echo=FALSE, include=FALSE}
icon_link(icon = "fas fa-images",
          text = "Slides",
          url = "")
icon_link(icon = "fas fa-download",
          text = "Download",
          url = "f")
```

## References

Dantas, G.P.M., Oliveira, L.R., Santos, A.M., Flores, M.D., De Melo, D.R., Simeone, A., González-Acuña, D., Luna-Jorquera, G., Le Bohec, C., Valdés-Velásquez, A., Cardeña, M., Morgante, J.S. & Vianna, J.A. 2019. Uncovering population structure in the Humboldt penguin (*Spheniscus humboldti*) along the Pacific coast at South America. PLoS One 14: 1–19. [Link](https://doi.org/10.1371/journal.pone.0215293)

De La Puente, S., Bussalleu, A., Cardeña, M., Valdés-Velásquez, A., Majluf, P. & Simeone, A. 2020. Humboldt Penguin (*Spheniscus humboldti*). Birds of the World 269–287. [Link](https://www.researchgate.net/publication/304626790_Humboldt_Penguin_Spheniscus_humboldti)

Herling, C., Culik, B.M. & Hennicke, J.C. 2005. Diet of the Humboldt penguin (*Spheniscus humboldti*) in northern and southern Chile. Mar. Biol. 147: 13–25. [Link](https://www.researchgate.net/publication/225603482_Diet_of_the_Humboldt_penguin_Spheniscus_humboldti_in_northern_and_southern_Chile)

Kaltenberg, A.M. & Benoit-Bird, K.J. 2009. Diel behavior of sardine and anchovy schools in the California Current System. Mar. Ecol. Prog. Ser. 394: 247–262. [Link](https://www.researchgate.net/publication/240809604_Diel_behavior_of_sardine_and_anchovy_schools_in_the_California_Current_System)

Quispe, R., Lerma, M., Serratosa, J. & Luna-Jorquera, G. 2020. Foraging ranges of Humboldt penguins *Spheniscus humboldti* from Tilgo Island: the critical need for protecting a unique marine habitat. Mar. Ornithol. 48: 205–208. [Link](https://www.marineornithology.org/PDF/48_2/48_2_205-208.pdf)

Taylor, J.C., Rand, P.S. & Jenkins, J. 2007. Swimming behavior of juvenile anchovies (Anchoa spp.) in an episodically hypoxic estuary: implications for individual energetics and trophic dynamics. Mar. Biol. 152: 939–957. [Link](https://doi.org/10.1007/s00227-007-0745-6)

Vianna, J. A. Cortes M., Ramos, B., Sallaberry, N., González-Acuña, D.,Dantas, G.P.M., Morgante, J., Simeone, A., Luna-Jorquera, G. Changes in abudance and distribution of Humboldt penguin *Spheniscus humboldti*. Mar. Ornithol. 42: 153-159. [Link](https://www.marineornithology.org/PDF/42_2/42_2_153-159.pdf)

### Credits

Navigation route at Tilgo Island by [Oceana](https://chile.oceana.org/blog/sitios-de-alimentacion-del-pinguino-de-humboldt-se-sobreponen-rutas-de-navegacion-del-Proyecto)

Humboldt penguin call by Carlos Pinto [xeno-canto](https://www.xeno-canto.org/species/Spheniscus-humboldti)



