---
title: "BLS7"
description: |
  Diel foraging behaviour of Humboldt penguins *Spheniscus humboldti* at Tilgo Island, Northern Chile.
preview: BLS_01.jpg
author:
  - name: Miriam Lerma
    url: {}
date: 09-23-2021
categories:
  - Seabird
  - presentation
  - English
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(distilltools)
```


# Intro

This post is dedicated to the poster and presentation for the Bio-Logging Symposium (https://www.bls7hawaii.com)

Diel foraging behaviour of Humboldt penguins *Spheniscus humboldti* at Tilgo Island, Northern Chile

Materials are underdevelopment 👩🏽‍🔧.

```{r icon-links1,echo=FALSE, include=TRUE}
icon_link(icon = "fas fa-images",
          text = "Preview poster",
          url = "https://raw.githubusercontent.com/MiriamLL/miriamlerma_/main/_posts/bls7/Poster_BSL7.pdf")
```

## Authors

**Miriam Lerma^1^**, Camila P. Villavicencio^2^, Nicolas Luna^3^, Matías Portflitt-Toro^3^, Juan Serratosa^3,4^, Guillermo Luna-Jorquera^3^, Stefan Garthe^1^, Rene Quispe^5,6^

1. Research and Technology Center (FTZ), University of Kiel, Hafentörn 1, 25761 Büsum, Germany. [Link](https://www.ftz.uni-kiel.de/en?set_language=en)
2. Instituto de Ecología y Biodiversidad, Departamento de Ciencias Ecológicas, Universidad de Chile, Las Palmeras 3425, Ñuñoa, Santiago, Chile. [Link](https://www.uchile.cl/portal/facultades-e-institutos/ciencias/160077/presentacion)
3. Millennium Nucleus for Ecology and Sustainable Management of Oceanic Islands - ESMOI, Larrondo 1281, Coquimbo, Chile. [Link](https://www.iniciativamilenio.cl/nucleoesmoi/)
4. BirdLife International, The David Attenborough Building, Pembroke Street, Cambridge, CB2 3QZ, UK
5. Departamento de Biología Marina, Facultad de Ciencias del Mar, Universidad Católica del Norte, Larrondo 1281, Coquimbo, Chile. [Link](https://www.ucn.cl/facultad-de-ciencias-del-mar/departamento-de-biologia-marina/)
6. Facultad de Ciencias Agropecuarias, Universidad Pedro de Valdivia, Av. Cuatro Esquinas 60, La Serena, Chile


## References

1. De La Puente, S., Bussalleu, A., Cardeña, M., Valdés-Velásquez, A., Majluf, P. & Simeone, A. 2020. Humboldt Penguin (*Spheniscus humboldti*). Birds of the World 269–287. [Link](https://www.researchgate.net/publication/304626790_Humboldt_Penguin_Spheniscus_humboldti)

2. Vianna, J. A. Cortes M., Ramos, B., Sallaberry, N., González-Acuña, D.,Dantas, G.P.M., Morgante, J., Simeone, A., Luna-Jorquera, G. Changes in abudance and distribution of Humboldt penguin *Spheniscus humboldti*. Mar. Ornithol. 42: 153-159. [Link](https://www.marineornithology.org/PDF/42_2/42_2_153-159.pdf)

3. Dantas, G.P.M., Oliveira, L.R., Santos, A.M., Flores, M.D., De Melo, D.R., Simeone, A., González-Acuña, D., Luna-Jorquera, G., Le Bohec, C., Valdés-Velásquez, A., Cardeña, M., Morgante, J.S. & Vianna, J.A. 2019. Uncovering population structure in the Humboldt penguin (*Spheniscus humboldti*) along the Pacific coast at South America. PLoS One 14: 1–19. [Link](https://doi.org/10.1371/journal.pone.0215293)

4. Quispe, R., Lerma, M., Serratosa, J. & Luna-Jorquera, G. 2020. Foraging ranges of Humboldt penguins *Spheniscus humboldti* from Tilgo Island: the critical need for protecting a unique marine habitat. Mar. Ornithol. 48: 205–208. [Link](https://www.marineornithology.org/PDF/48_2/48_2_205-208.pdf)

5. Navigation route of proposed industrial project at Tilgo Island by [Oceana](https://chile.oceana.org/blog/sitios-de-alimentacion-del-pinguino-de-humboldt-se-sobreponen-rutas-de-navegacion-del-Proyecto)

6. Herling, C., Culik, B.M. & Hennicke, J.C. 2005. Diet of the Humboldt penguin (*Spheniscus humboldti*) in northern and southern Chile. Mar. Biol. 147: 13–25. [Link](https://www.researchgate.net/publication/225603482_Diet_of_the_Humboldt_penguin_Spheniscus_humboldti_in_northern_and_southern_Chile)

7. Kaltenberg, A.M. & Benoit-Bird, K.J. 2009. Diel behavior of sardine and anchovy schools in the California Current System. Mar. Ecol. Prog. Ser. 394: 247–262. [Link](https://www.researchgate.net/publication/240809604_Diel_behavior_of_sardine_and_anchovy_schools_in_the_California_Current_System)

8. Taylor, J.C., Rand, P.S. & Jenkins, J. 2007. Swimming behavior of juvenile anchovies (Anchoa spp.) in an episodically hypoxic estuary: implications for individual energetics and trophic dynamics. Mar. Biol. 152: 939–957. [Link](https://doi.org/10.1007/s00227-007-0745-6)

9. Humboldt penguin call by Carlos Pinto [xeno-canto](https://www.xeno-canto.org/species/Spheniscus-humboldti)

## Acknowledgments

This study was financed by Fondecyt (grant no. 3170936 to RQ and postdoctorate grant no. 3160679 to CPV) and by CONACyT-INAPI (grant no. 411876 to ML). 

We thank Esther Rickert for the transportation of tracking devices, and Henri Kröling for useful discussions. 

This study was carried out in accordance with the recommendations and permit of the Subsecretaria de Pesca y Acuicultura de Chile (resolution N°3367/2017). 
