---
title: "Plot bathymetry in R"
description: |
  Create a map with bathymetry information in R.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-01-10
preview: blog41_portadas.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---

# Intro

Bathymetry gives us information on the water depth around an area.

This post includes:  
- Download raster data   
- Read and subset raster data  
- Plot raster data  

# Download raster data

GEBCO: General Bathymetry Chart of the Oceans provides information from bathymetry in the ocean.

To download [visit webpage](https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area)

Go to:  
- Download data for user-defined areas  
- [Use the application](https://download.gebco.net/)  
- Add your coordinates -here I use 1 to 10 and 50 to 60-  
- Add to basket and download

# Load raster data

Once you had download the information, give the direction of the file. For example:

```{r, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"this_directory")
this_file<-paste0(this_folder,"this_file")
```

```{r, include=FALSE, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2025-01-15-bathymetry")
list.files(this_folder)
```

```{r, include=FALSE, eval=FALSE}
this_file<-paste0(this_folder,"/gebco_2024_n60.0_s50.0_w1.0_e9.0.tif")
this_file
```

Load the [terra package](https://datacarpentry.org/r-raster-vector-geospatial/01-raster-structure.html) for reading raster data

```{r}
library(terra)
```

The function **rast** helps to read raster data - replacing package raster

```{r, eval=FALSE}
Bath_file<-rast(this_file)
```

Change to data frame

```{r, eval=FALSE}
Bath_dataframe <- as.data.frame(Bath_file, xy = TRUE)
```

Alternatively, use the test information included in the package **GermanNorthSea**

```{r}
##devtools::install_github("MiriamLL/GermanNorthSea")
library(GermanNorthSea)
```

```{r}
Bath_dataframe<-GermanNorthSea::German_bath
```

Load the package **tidyverse** 

```{r}
library(tidyverse)
```

Use **filter** to subset your data 

```{r}
Bath_dataframe_sub <-Bath_dataframe  %>%
  filter(x > 2 & x < 10)%>%
  filter(y > 52 & y < 57)%>%
  rename(Bathymetry=3) %>%
  filter(Bathymetry < 10)
```


# Plot

Load the package **ggplot2** 

```{r}
library(ggplot2)
```

Plot your data using **geom_raster**

```{r, fig.width=4, fig.height=4}
ggplot() +
  geom_raster(data = Bath_dataframe_sub , aes(x = x, y = y, fill = Bathymetry)) +
  scale_fill_viridis_c(option = "mako")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))

```


Load the package **sf**

```{r}
library(sf)
```
Use the function **st_transform** to convert to the same CRS

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```

Add the land to the plot using **geom_sf**

```{r, fig.width=4, fig.height=4}
Plot_bath<-ggplot() +
  geom_raster(data = Bath_dataframe_sub , aes(x = x, y = y, fill = Bathymetry)) +
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "mako")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_bath
```

More details about how to add features to a map [here](https://www.miriam-lerma.com/posts/2023-03-04-mapping-in-r/)

```{r, echo=FALSE, eval=FALSE}
ggsave(Plot_bath,
       filename = paste0(this_folder,"/blog41.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

- [Package GermanNorthSea](https://github.com/MiriamLL/GermanNorthSea)
- [Package terra](https://rspatial.github.io/terra/) 
- [Package sf](https://r-spatial.github.io/sf/)
- [GEBCO download](https://download.gebco.net/)

