---
title: "Fishing effort in the German North Sea"
description: |
  Create a map with sst information from Global fishing watch in R. 
author:
  - name: Miriam Lerma
    url: {}
date: 2025-04-15
preview: Y2025_apr_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---


# Data

[Global fishing watch](https://globalfishingwatch.org/datasets-and-code/)

To use the data, you need to register (is free of use).

Login using your email.

[Fishing effort](https://globalfishingwatch.org/map?access-token=wi1_Ffdm-AjT1k_urFVmOPTMqMXrKQ)




# Data

The Global Fishing Watch map is the first open-access online tool for visualization and analysis of vessel-based human activity at sea. Anyone with an internet connection can access the map to monitor global fishing activity from 2012 to the present for more than 65,000 commercial fishing vessels that are responsible for a significant part of global seafood catch.

We use data that is broadcast using the automatic identification system (AIS) and collected via satellites and terrestrial receivers. We then combine this information with vessel monitoring system data provided by our partner countries. We apply our fishing detection algorithm to determine “apparent fishing effort” based on changes in vessel speed and direction. The heat map grid cell colors show how much fishing happened in that area, allowing for precise comparison.

I want data of apparent fishing effort (AIC)

Mz first step is to limit the data to April and May of 2018.

[Video](https://www.youtube.com/watch?v=iWkgapKbJWg)

```{r}

```

# Subset


```{r}

```

# Plot

```{r}
German_land<-st_transform(GermanNorthSea::German_land, 4326)
```


Use **ggplot** to create your plot. 

```{r, fig.width=4, fig.height=4}
Plot_distance<-ggplot() +
  geom_raster(data = DistCoast_dataframe_sub, aes(x = x, y = y, fill = Dist)) +
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  scale_fill_viridis_c(option = "rocket")+
  theme_void()+
  theme(legend.position='bottom')+
  xlab('Longitude')+ylab('Latitude')+
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
Plot_distance+
  guides(fill=guide_legend(title="Distance to coast"))
```

# Export

```{r, echo=FALSE, eval=FALSE}
library(here)
here()
this_folder<-paste0(here(),"/_posts/2024-10-15-distancetocoast")
ggsave(Plot_distance,
       filename = paste0(this_folder,"/blog42.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```








# Further read

[Working with public data](https://globalfishingwatch.org/data/downloadable-public-data-in-r/#)
[GitHub with outdate instructions](https://github.com/GlobalFishingWatch/paper-global-footprint-of-fisheries/blob/master/data_documentation/fishing_effort_byvessel.md)
[](https://www.science.org/doi/10.1126/science.aao5646)