---
title: "Raster Dist Colony"
description: |
  Create a raster using information from distance to colonies. 
author:
  - name: Miriam Lerma
    url: {}
date: 2025-04-10
preview: blog44_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - raster
  - English
  - Y2025
---

# Intro

In this post, we would create a raster using distance from other colonies.

# Data

The data corresponds to colonies of gannets.

```{r}
colony_lat=c(55.25,56.78,56.08,54.15,51.59,
             53.80,59.54,58.29,60.13,51.73,
             52.12,54.18,60.71,53.40,53.49,
             49.70,51.78,62.10,60.14,49.72,
             48.90,54.67,57.81,59.10,59.08,
             59.02,57.69,59.29)
colony_long=c(-5.12,-7.64,-2.64,-0.17,-10.27,
              -10.00,-1.63,-7.59,-2.07,-5.48,
              -6.62,7.89,-0.88,-6.06,-6.02,
              -2.24,-10.51,-7.68,-1.02,-2.29,
              -3.44,-4.70,-8.56,-6.16,-4.41,
              -4.51,-2.29,-2.98)
```

Use the data to create a matrix using **cbind**

```{r}
this_colonies<-cbind(colony_long,colony_lat)
```

Transform to vector using the package **terra**

```{r}
library(terra)
```

Provide CRS

```{r}
this_vector <- vect(this_colonies, 
                    crs="+proj=longlat +datum=WGS84")
```

Identify limits

```{r, include=FALSE}
min(colony_lat)
max(colony_lat)
min(colony_long)
max(colony_long)
```

Create raster using the function **rast**  
This function creates a raster from scratch  
Use the same crs as in the vector  

```{r}
this_raster <- rast(ncols=200, nrows=200, 
                    crs="+proj=longlat +datum=WGS84",
                    xmin = -15, xmax = 15, 
                    ymin = 40, ymax = 70)
this_raster[500] <- 1
```

# Distance

Calculate distance using the function **distance** from the package terra  

```{r}
this_calculation <- distance(this_raster, this_vector)
```

To access the data use **as.data.frame**  

```{r}
this_layers<-as.data.frame(this_calculation, xy = TRUE)
```

Transform from meters to kilometers

```{r}
this_layers$kms<-this_layers$lyr.1/1000
```

Check ranges

```{r}
mean(this_layers$kms)
range(this_layers$kms)
```

Plot using ggplot

```{r}
library(ggplot2)
```

Use the **geom_raster**  

```{r, fig.height=6, fig.width=6}
ggplot(data = this_layers) +
    geom_raster(aes(x = x, y = y, fill = kms)) +
    scale_fill_viridis_c() +
    theme_void() +
    theme(legend.position = "bottom")
```

Create data frame with the information from the colonies  

```{r}
colonies_sub<-data.frame(
  name=c("Ailsa Craig","Barra Head","Bass Rock","Bempton Cliff","Bull Rock","Clare Island", "Fair Isle", "Flannan Isles","Foula","Grassholm Island","Great Saltee", "Helgoland" , "Hermaness", "Ireland’s Eye", "Lambay", "Les Etac/Alderrney","Little Skellig","Mykinesholmur","Noss" ,"Ortac","Rouzic Island","Scar Rocks","St. Kilda","Sula sgeir","Sule Skerry","Sule Stack","Troup Head","Westray"),
colony_lat=colony_lat,
colony_long=colony_long)
```

Include in the plot using **geom_point**  

```{r, fig.height=6, fig.width=6}
ggplot(data = this_layers) +
    geom_raster(aes(x = x, y = y, fill = kms)) +
    scale_fill_viridis_c() +
    theme_void() +
    theme(legend.position = "bottom")+
  geom_point(data=colonies_sub,aes(x=colony_long,y=colony_lat),color='white')
```

Add land for a better reference

```{r}
library(sf)
```

Transform the CRS

```{r}
Land_europe<-GermanNorthSea::German_land
Land_europe<-st_transform(Land_europe,4326)
```

Include as **geom_sf** 

```{r, fig.height=6, fig.width=6}
ggplot(data = this_layers) +
  geom_raster(aes(x = x, y = y, fill = kms)) +
  scale_fill_viridis_c() +
  
  geom_sf(data=Land_europe, colour = '#283618', fill = 'transparent') +
  geom_point(data=colonies_sub,aes(x=colony_long,y=colony_lat),color='#c1121f')+
  
  theme_bw()+
  
  theme(
  panel.background = element_rect(fill = '#edf2f4'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),legend.position='none',
  panel.border = element_rect(colour = "black", fill=NA, size=1.5),
  
  axis.text.x = element_text(color='#283618',size=10,vjust = 10),
  axis.text.y = element_text(color='#283618',size=10,margin = margin(0,-1.00,0,1, unit = 'cm')),
  axis.title = element_blank(),
  axis.ticks.length=unit(-0.20, "cm"))+
  
  coord_sf(xlim = c(-12, 10),ylim = c(48, 63))+
  scale_x_continuous(breaks = c(-10,-5,0,5,10,20,30),labels = function(x) paste0(x, '\u00B0')) +
  scale_y_continuous(breaks = c(45,50,55,60,65,70,75),labels = function(x) paste0(x, '\u00B0')) +
  
  theme(legend.position='bottom')+
  
  NULL
```

# Crop

Crop the raster to have a more adequate legend

```{r}
crop_extention<-ext(-15, 12, 45, 65)
crop_calculation<-crop(this_calculation,crop_extention)
```

Transform to data frame to acces the info

```{r}
crop_df<-as.data.frame(crop_calculation, xy = TRUE)
crop_df$kms<-crop_df$lyr.1/1000
```

```{r}
mean(crop_df$kms)
range(crop_df$kms)
```

Substitute the data used

```{r, fig.height=6, fig.width=6}
ggplot(data = crop_df) +
  geom_raster(aes(x = x, y = y, fill = kms)) +
  scale_fill_viridis_c() +
  
  geom_sf(data=Land_europe, colour = '#283618', fill = 'transparent') +
  geom_point(data=colonies_sub,aes(x=colony_long,y=colony_lat),color='#c1121f')+
  
  theme_bw()+
  
  theme(
  panel.background = element_rect(fill = '#edf2f4'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),legend.position='none',
  panel.border = element_rect(colour = "black", fill=NA, size=1.5),
  
  axis.text.x = element_text(color='#283618',size=10,vjust = 10),
  axis.text.y = element_text(color='#283618',size=10,margin = margin(0,-1.00,0,1, unit = 'cm')),
  axis.title = element_blank(),
  axis.ticks.length=unit(-0.20, "cm"))+
  
  coord_sf(xlim = c(-12, 10),ylim = c(48, 63))+
  scale_x_continuous(breaks = c(-10,-5,0,5,10,20,30),labels = function(x) paste0(x, '\u00B0')) +
  scale_y_continuous(breaks = c(45,50,55,60,65,70,75),labels = function(x) paste0(x, '\u00B0')) +
  
  theme(legend.position='bottom')+
  
  NULL
```

# Mask

To remove the land information use the function **mask** 

```{r}
mask_calculation <- mask(crop_calculation, Land_europe, inverse = TRUE)
```

Again check the data frame

```{r}
mask_df<-as.data.frame(mask_calculation, xy = TRUE)
mask_df$kms<-mask_df$lyr.1/1000
```

```{r}
mean(mask_df$kms)
range(mask_df$kms)
```

Substitute the data, now it would not include the land. 

```{r, fig.height=6, fig.width=6}
ggplot(data = mask_df) +
  geom_raster(aes(x = x, y = y, fill = kms)) +
  scale_fill_viridis_c() +
  
  #geom_sf(data=Land_europe, colour = '#283618', fill = 'transparent') +
  geom_point(data=colonies_sub,aes(x=colony_long,y=colony_lat),color='#c1121f')+
  
  theme_bw()+
  
  theme(
  panel.background = element_rect(fill = '#edf2f4'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),legend.position='none',
  panel.border = element_rect(colour = "black", fill=NA, size=1.5),
  
  axis.text.x = element_text(color='#283618',size=10,vjust = 10),
  axis.text.y = element_text(color='#283618',size=10,margin = margin(0,-1.00,0,1, unit = 'cm')),
  axis.title = element_blank(),
  axis.ticks.length=unit(-0.20, "cm"))+
  
  coord_sf(xlim = c(-12, 10),ylim = c(48, 63))+
  scale_x_continuous(breaks = c(-10,-5,0,5,10,20,30),labels = function(x) paste0(x, '\u00B0')) +
  scale_y_continuous(breaks = c(45,50,55,60,65,70,75),labels = function(x) paste0(x, '\u00B0')) +
  
  theme(legend.position='bottom')+
  
  NULL
```




# Export

Select export folder

```{r, eval=TRUE, echo=FALSE}
library(here)
this_folder<-paste0(here(),"/_posts/2025-07-15-distancecolonyraster")
```

```{r, eval=FALSE}
this_folder
```

Select export raster file

```{r eval=FALSE}
writeRaster(this_calculation, filename=file.path(this_folder, "distance_colonies_raster.tif"))
```

```{r, eval=FALSE, echo=FALSE}
library(tidyverse)
write_csv(colonies_sub,paste0(this_folder,"/colonies.txt"))
```

```{r, eval=FALSE, echo=FALSE}
Plot_crop<-ggplot(data = crop_df) +
  geom_raster(aes(x = x, y = y, fill = kms)) +
  scale_fill_viridis_c() +
  
  geom_sf(data=Land_europe, colour = '#283618', fill = 'transparent') +
  geom_point(data=colonies_sub,aes(x=colony_long,y=colony_lat),color='#c1121f')+
  
  theme_bw()+
  
  theme(
  panel.background = element_rect(fill = '#edf2f4'),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),legend.position='none',
  panel.border = element_rect(colour = "black", fill=NA, size=1.5),
  
  axis.text.x = element_text(color='#283618',size=10,vjust = 10),
  axis.text.y = element_text(color='#283618',size=10,margin = margin(0,-1.00,0,1, unit = 'cm')),
  axis.title = element_blank(),
  axis.ticks.length=unit(-0.20, "cm"))+
  
  coord_sf(xlim = c(-12, 10),ylim = c(48, 63))+
  scale_x_continuous(breaks = c(-10,-5,0,5,10,20,30),labels = function(x) paste0(x, '\u00B0')) +
  scale_y_continuous(breaks = c(45,50,55,60,65,70,75),labels = function(x) paste0(x, '\u00B0')) +
  
  theme(legend.position='none')+
  
  NULL
Plot_crop
```

```{r, echo=FALSE, eval=FALSE}
library(here)
here()
ggsave(Plot_crop,
       filename = paste0(this_folder,"/blog57.jpg"),
       width = 4, 
       height = 4, 
       units = "in",
       dpi = 300)
```

# Further reading

- [distance](https://github.com/rspatial/terra/issues/567)
- [crop](https://rspatial.github.io/terra/reference/crop.html?q=crop#null)
- [mask](https://rspatial.github.io/terra/reference/mask.html?q=mask#null)



Other uses could be use include distance to the nearest boat, for gulls for example.


# End of document