---
title: "Habitat use"
description: |
  How to assign habitat use to animal locations.
author:
  - name: Miriam Lerma
    url: {}
date: 2022-04-07
preview: blog10b.jpg
categories:
  - R
  - Tracking
  - English
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    self_contained: false
---


# Intro

This post is about how to classify locations on your tracking data based on an polygon with geographical information.

This is useful for example when you want to know:  
- If the animal is using a specific habitats, and  
- To calculate the percentage of use of a specific habitat  
See examples on the recommended literature at the bottom of this post.

In this post we will classify tracking locations as **land** vs **not land**.

# Data 📖

To do this exercise, load data from the package 'sula'. <br>
For accessing the data, you need to have the package installed. 

To install:

```{r, eval=FALSE}
#devtools::install_github("MiriamLL/sula")
```

```{r}
library(sula)
```

The data is from 10 tracked individuals.

```{r}
my_locs<-(GPS_raw)
```


Check how many locations I have per individual

```{r, warnings=FALSE}
library(tidyverse)
```

```{r}
my_locs %>%
  drop_na(IDs)%>%
  group_by(IDs)%>%
  tally()
```


# Points ⏩

To transform your GPS locations to spatial data, use the functions from the package **sp**.   
It would turn your data frame into a **SpatialPointsDataFrame**.

```{r}
my_locs <- as.data.frame(my_locs)
my_points <-my_locs
sp::coordinates(my_points) <- ~Longitude + Latitude
sp::proj4string(my_points) = sp::CRS("+init=epsg:4326")
```

# Polygon 🗺️

For the polygon, you can download shapefiles according to your interest and area of study. 

Some sources of country polygons are:  
- [DIVA-GIS](https://www.diva-gis.org/datadown)

But there are many shapefile sources per country,  
- For **Chile**: several exammples [are summarized here](https://sporella.xyz/2020/12/01/30daymap/)  
- For **Mexico**: IDEA [Infraestructura de Datos Espaciales Abiertos](https://www.gits.igg.unam.mx/idea/descarga)  
- For **Germany**: [BSH-GeoSeaPortal](https://www.bsh.de/EN/DATA/GeoSeaPortal/geoseaportal_node.html;jsessionid=8CAD8F86337BF25B4A2B2A7F13C0CDC8.live11311)  

... just to mention a few.


Because the tracking data from the example was collected in Chile, we will use the country shapefile of Chile downloaded from the [DIVA-GIS](https://www.diva-gis.org/datadown) portal.

```{r, include=FALSE}
library(here)
```

```{r, echo=FALSE}
ShapefilesDirectory<-here('_posts/2022-04-07-land-or-not')
```

The package **rgdal** has the function **readOGR** which allows us to load the polygon into R. 

```{r}
library(rgdal)
```

The polygon will be read as **SpatialPolygonDataFrame**.

```{r, warning=FALSE, message=FALSE, eval=FALSE}
my_polygon<-readOGR(file.path(ShapefilesDirectory,'CHL_adm0.shp')) 
```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
my_polygon<-readOGR(file.path(ShapefilesDirectory,'CHL_adm0.shp')) 
```

# Over 🏞️

The function **over** checks which of the points (my_locs) fall inside the polygon (my_polygon).

Be cautious, the points and the polygon should be at the same CRS.

```{r}
polygon_proj <- proj4string(my_polygon)
```

```{r}
proj4string(my_points) <- polygon_proj
```

The function **over** will return a data frame with the information where the points and the polygon intersect.

```{r}
Landornot_over<- over(my_points, my_polygon)
```

It depends on the information of your polygon, but in this particular example, in the column NAME_0, if the points fall inside the polygon, it keeps 'Chile' otherwise returns an **NA**.  

```{r}
unique(Landornot_over$NAME_0)
```

We can add this information back to our locations to know which of the locations fall inside the territorial land limits or not.

```{r}
my_locs$landornot <- Landornot_over$NAME_0
```

To make it easier to remember you can replace Chile and 0 for the data available in your polygon. 

```{r}
my_locs$landornot[is.na(my_locs$landornot)] <- 0
```

```{r}
my_locs<-my_locs %>%
  mutate(landornot=
           case_when(
             landornot == 0 ~ 'notland',
             landornot != 0 ~ 'land',
             TRUE ~ "U"))
```

You can also export this data frame to check in a GIS program.

```{r, eval=FALSE}
write_csv(my_locs,file =paste0(here("1Data"),'/my_locs.csv'))
```

# Percentages 🥧

Now using functions from the package **tidyverse** you can quantify the numer of locations that the animal was at land or not.

```{r}
Habitatuse<-my_locs %>%
  drop_na(IDs)%>%
  group_by(IDs,landornot)%>%
  count()%>%
  pivot_wider(names_from = landornot, values_from = n)
```

... and also calculate the percentage of time using one habitat or the other.

```{r}
Habitatuse$total<-Habitatuse$land+Habitatuse$notland
Habitatuse$prop_land<-Habitatuse$land*100/Habitatuse$total
Habitatuse$prop_notland<-Habitatuse$notland*100/Habitatuse$total
Habitatuse$prop<-Habitatuse$prop_land+Habitatuse$prop_notland
```

# Plot 🖌️

## Map

Plotting the data will help to confirm if the classification is correct.  

```{r, fig.height=5, fig.width=4}
ggplot()+
  geom_polygon(data=my_polygon, aes(x=long, y=lat, group=group), color='black',fill='grey')+
  geom_point(data=my_locs, aes(x=Longitude,y=Latitude,color=landornot))+
  scale_color_manual(values=c('#9b2226','#005f73'))+
  theme_bw()+
  theme(legend.position = 'top')+
  labs(x = "Longitude", y="Latitude",color='')+
  xlim(-110, -108)+
  ylim(-28.5, -26.5)
```

To customize your plot, you can change many arguments, for example to change the **theme_bw** [here are other options](https://ggplot2.tidyverse.org/reference/ggtheme.html).  
**Do not forget** to adjust your xlim and ylim to your coordinates. 


## Percentages

You can also plot the percentage of habitat use per individual.

To do this, preparing the information in a *long* format will make it easier to plot. 

```{r}
Habitatuse_prop<-Habitatuse %>%
  pivot_longer(c(prop_land,prop_notland),
               names_to = "habitat_use",
               values_to = "prop_use" )
```

```{r}
ggplot(Habitatuse_prop, aes(fill=habitat_use, y=prop_use, x=IDs)) + 
  geom_bar(position="fill", stat="identity")+
  scale_fill_manual(values=c('#9b2226','#005f73'))+
  theme_bw()+
  theme(legend.position = 'top')+
  scale_y_continuous(expand = c(0,0)) 
```

# Recommended literature 👩🏽‍🏫

- Function **over**:  
[Geometry overlays](https://cran.r-project.org/web/packages/sp/vignettes/over.pdf)

- Examples in the literature:  
[Quantify habitats visited](https://onlinelibrary.wiley.com/doi/10.1111/ibi.12917)  
[Individual specialization in an ecological winner](https://www.int-res.com/articles/theme/m578p243.pdf)  
[Tracking positions inside the protected areas](https://www.researchgate.net/publication/258062415_Correction_Do_Power_Lines_and_Protected_Areas_Present_a_Catch-22_Situation_for_Cape_Vultures_Gyps_coprotheres)

If you can think on another sources of geographical data please do not hesitate to write me to include it on the list. 

For more details or suggestions, or just to say hi you can also always write me an [email](miriamjlerma@gmail.com)
