---
title: "shipping traffic"
description: |
 Create a map including shipping traffic information.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-06-02
preview: shipping_portada.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2025
---

# Intro

The Vessel Density maps in the EU are created since the 2019 by Cogea for the European Marine Observation and Data Network [EMODnet](https://emodnet.ec.europa.eu/geoviewer/#!/). The dataset is updated every year and is available for viewing and download on EMODnet Human Activities web portal (https://emodnet.ec.europa.eu/en/human-activities).   
The maps are based on AIS data yearly purchased from Collecte Localisation Satellites (CLS) and ORBCOMM. The maps, GeoTIFF format, show shipping density in 1x1km cells of a grid covering all EU waters and some neighbouring areas.  
Density is expressed as hours per square kilometre per month. The following ship types are available:0 Other, 1 Fishing, 2 Service, 3 Dredging or underwater ops, 4 Sailing, 5 Pleasure Craft, 6 High speed craft, 7 Tug and towing, 8 Passenger, 9 Cargo, 10 Tanker, 11 Military and Law Enforcement, 12 Unknown and All ship types.  
Data are available by month of year. Yearly averages are also available.

## Data

To download go to [EMODnet Map Viewer](https://emodnet.ec.europa.eu/geoviewer/) > Catalogue > EMODnet Human Activities > Vessel density > Monthly totals 2017-2023 > All Types > Click on the "i" icon > Select Download.

A zip file named "EMODnet_HA_Vessel_Density_all_2017-2023 would be downloaded. For the period of December 2023, the raster information is under **vesseldensity_all_20231201.tif**.

The original file includes information from a larger area, to focus on a specific, the function **crop** can be used.   [Crop](https://search.r-project.org/CRAN/refmans/raster/html/crop.html) returns a geographic subset of an object as specified by an Extent object (or object from which an extent object can be extracted/created). If x is a Raster* object, the Extent is aligned to x. Areas included in y but outside the extent of x are ignored (see extend if you want a larger area). 

```{r, echo=FALSE, eval=FALSE, message=FALSE}
# Read original file 
your_folder<-paste0(here::here(),"/_posts/2025-06-12-shippingtraffic")
your_file<-paste0(your_folder,"/vesseldensity_all_20231201.tif")
ShippingTraffic<-terra::rast(your_file)
# Plot
ggplot2::ggplot() +
  tidyterra::geom_spatraster(data = ShippingTraffic) +
  NULL
# Crop
library(raster)
polygon_crop <- as(extent(3790000,4250000, 3350000,3680000), 'SpatialPolygons')
crs(polygon_crop) <- "+proj=longlat +datum=WGS84 +no_defs"
ShippingTraffic_crop<- crop(ShippingTraffic, polygon_crop)
# Export
writeRaster(ShippingTraffic_crop, "ShippingTraffic_crop.tif")
```

## Load

To download test data in tif format [click here](https://github.com/MiriamLL/TestData/blob/main/ShippingTraffic_crop.tif).  

Use the package **terra** to use the function **rast**.  

```{r, message=FALSE}
library(terra)
```

The function **rast** reads your file as a formal class SpatRaster. Here I am using the file from the repository. 

```{r}
ShippingTraffic_tif<-'https://github.com/MiriamLL/TestData/raw/refs/heads/main/ShippingTraffic_crop.tif'
```

```{r}
ShippingTraffic<-rast(ShippingTraffic_tif)
```

```{r, eval=FALSE}
crs(ShippingTraffic)
```

```{r, fig.height=5, fig.width=5}
ggplot2::ggplot() +
  tidyterra::geom_spatraster(data = ShippingTraffic) +
  NULL
```

## Crop and Mask

Select only values inside the EEZ

```{r}
German_EEZ<-GermanNorthSea::German_EEZ
polygon_crop <- German_EEZ
```

```{r, fig.width=5, fig.height=5}
ShippingTraffic_crop <- crop(ShippingTraffic, polygon_crop)
ShippingTraffic_mask <- mask(x= ShippingTraffic_crop, mask=polygon_crop)
plot(ShippingTraffic_mask)
```

## Data frame

```{r, message=FALSE}
library(tidyverse)
```

```{r}
ShippingTraffic_df <- as.data.frame(ShippingTraffic_mask, xy = TRUE)
beepr::beep(sound=1)
```

The function rename, allows to change the name of the column. The density information is in the third column.

```{r}
ShippingTraffic_df<-ShippingTraffic_df %>%
  rename(vessel_density=3)
```

Check if the density values are plausible.

```{r}
range(ShippingTraffic_df$vessel_density)
```

Values distribution

```{r, fig.height=5, fig.width=5}
hist(ShippingTraffic_df$vessel_density)
```

```{r, echo=TRUE, eval=FALSE}
ShippingTraffic_class<-ShippingTraffic_df %>%
 mutate(
    density_class = case_when(
      vessel_density <= 0  ~ "class_A_0",
      vessel_density > 0 &  vessel_density < 1 ~ "class_A_0_1",
      vessel_density > 1 &  vessel_density < 2 ~ "class_A_1_2",
      vessel_density > 2 &  vessel_density < 3 ~ "class_A_2_3",
      vessel_density > 3 &  vessel_density < 4 ~ "class_A_3_4",
      vessel_density > 4 &  vessel_density < 5 ~ "class_A_4_5",
      vessel_density > 5 &  vessel_density < 6 ~ "class_A_5_6",
      vessel_density > 6 &  vessel_density < 7 ~ "class_A_6_7",
      vessel_density > 7 &  vessel_density < 8 ~ "class_A_7_8",
      vessel_density > 8 &  vessel_density < 9 ~ "class_A_8_9",
      vessel_density > 9 &  vessel_density < 10 ~ "class_A_9_10",
      
      vessel_density >= 10 &  vessel_density < 100 ~ "class_B_10_100",
      vessel_density >= 100 &  vessel_density < 1000 ~ "class_C_100_1000",
      vessel_density >= 1000 &  vessel_density < 10000 ~ "class_D_1000_10000",
      vessel_density >= 10000 &  vessel_density < 100000 ~ "class_E_10000_100000",
      TRUE  ~ "unknown"
    ))
```

The majority of values are between 0 and 1

```{r, echo=FALSE, eval=FALSE}
ShippingTraffic_class %>%
  group_by(density_class)%>%
  tally()
```

## Select area

Use the function **geom_spatraster** to plot the vessel density data.  

```{r}
library(tidyterra)
```

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask)
```

## Land

Add land for reference using **geom_sf** and the package **GermanNorthSea**.

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```


## Theme

Change background colors using the arguments on theme.

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  theme_void()+
    theme(panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, size=1.5))
```


## Gradient colors

Low traffic 

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+

   
  theme_void()+
    theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
    
  scale_fill_gradientn(name='Low Vessel density \n Dec 2023',
                      na.value = "transparent", 
                      colours = palette_yellow<-c("#ffff3f","#eeef20","#dddf00","#d4d700","#bfd200","#aacc00","#80b918","#55a630","#2b9348","#007f5f"),
                      limits = c(0.00001,10),
                      breaks = c(2,4,6,8,10),
                      )
```

Medium traffic

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
    scale_fill_gradientn(name='Medium Vessel density \n Dec 2023',
                      na.value = "transparent", 
                      colours = palette_green<-c("#d8f3dc","#b7e4c7","#95d5b2","#74c69d","#52b788","#40916c","#2d6a4f","#1b4332","#081c15"),
                      limits = c(5,100),
                      breaks = c(5,20,40,60,80,100),
                      )
  
```

High traffic

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        )+
  scale_fill_gradientn(name='High Vessel density \n Dec 2023 \n (hrs per km2)',
                      na.value = "transparent", 
                      colours = palette_red<-c("#e01e37","#da1e37","#c71f37","#bd1f36","#b21e35","#a71e34","#a11d33","#85182a","#6e1423","#641220"),
                      limits = c(100,30000),
                      breaks = c(0,1000,10000,20000,30000),
                      )
```

## Palette

```{r}
your_palette<-c("transparent", "#d8f3dc",
                "#b7e4c7","#95d5b2","#74c69d","#52b788","#aacc00",
                "#bfd200","#d4d700","#dddf00","#eeef20","#ffff3f", 
                "#e01e37","#da1e37","#c71f37", "#bd1f36","#b21e35",
                "#a71e34","#a11d33","#85182a","#6e1423","#641220")
```

Use **scale_fill_gradientn** and **rescale** to adjust the palette and the legend

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0.05,0.1,0.2,0.5,2,5,10,20,100,1000),
                      values = scales::rescale(c(0,0.01,0.05,0.1,0.2,0.5,2,5,10,20,100,1000)), 
                      guide = "legend"
                      )
```

Use **labels** to make the legend more legible

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0.05,0.1,0.2,0.5,2,5,10,20,100,1000),
                      values = scales::rescale(c(0,0.01,0.05,0.1,0.2,0.5,2,5,10,20,100,1000)), 
                      guide = "legend",
                      labels = c("0 - 0.05","> 0.1","> 0.2","> 0.5","> 2","> 5","> 10","> 20", "> 100","> 1,000") )
```


Include the legend inside the plot and change the theme, also change panel.background

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  scale_fill_gradientn(name="Vessel density \n inside Germany's EEZ \n Dec 2023 \n hrs per km2",
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0.05,0.1,0.2,0.5,2,5,10,20,100,1000),
                      values = scales::rescale(c(0,0.01,0.05,0.1,0.2,0.5,2,5,10,20,100,1000)), 
                      guide = "legend",
                      labels = c("0 - 0.05","> 0.1","> 0.2","> 0.5","> 2","> 5","> 10","> 20", "> 100","> 1,000"))+
  
  theme(legend.position = c(0.15,0.50),
          legend.background = element_rect(colour = FALSE, fill = FALSE),
          legend.title=element_text(color='black',size=16),
          legend.text=element_text(color='black',size=12),
          legend.key = element_rect(colour = 'transparent', fill = 'transparent'),
          legend.key.height = unit(8, "mm"))
```


I additionally added protected areas. 

```{r, fig.height=10, fig.width=10}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='black',fill='transparent',alpha=0.1, size = 1, linewidth=1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black')+
  
  geom_sf(data = GermanNorthSea::German_SCA, colour = 'black', fill = "transparent", linewidth=1, linetype = "dashed")+
  geom_sf(data = GermanNorthSea::German_natura, colour = 'black', fill = "transparent", linewidth=1, linetype = "dashed")+
  
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  scale_fill_gradientn(name="Vessel density \n Inside Germany's EEZ \n Dec 2023 \n hrs per km2",
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0.05,0.1,0.2,0.5,2,5,10,20,100,1000),
                      values = scales::rescale(c(0,0.01,0.05,0.1,0.2,0.5,2,5,10,20,100,1000)), 
                      guide = "legend",
                      labels = c("0 - 0.05","> 0.1","> 0.2","> 0.5","> 2","> 5","> 10","> 20", "> 100","> 1,000"))+
  
  theme(legend.position = c(0.15,0.50),
          legend.background = element_rect(colour = FALSE, fill = FALSE),
          legend.title=element_text(color='black',size=16),
          legend.text=element_text(color='black',size=12),
          legend.key = element_rect(colour = 'transparent', fill = 'transparent'),
          legend.key.height = unit(8, "mm"))
```

```{r, echo=FALSE, eval=FALSE}
ggsave(ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme_void()+
    theme(
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0.05,0.1,0.2,0.5,2,5,10,20,100,1000),
                      values = scales::rescale(c(0,0.01,0.05,0.1,0.2,0.5,2,5,10,20,100,1000)), 
                      guide = "legend",
                      labels = c("0 - 0.05","> 0.1","> 0.2","> 0.5","> 2","> 5","> 10","> 20", "> 100","> 1,000"))+
  
  theme(legend.position = c(0.15,0.50),
          legend.background = element_rect(colour = FALSE, fill = FALSE),
          legend.title=element_text(color='white',size=16),
          legend.text=element_text(color='white',size=12),
          legend.key = element_rect(colour = 'transparent', fill = 'transparent'),
          legend.key.height = unit(8, "mm")),
  
       filename = paste0(your_folder,"/VesselDensity_GermanNorthSea.jpg"),
       width = 10, 
       height =10, 
       units = "in",
       dpi = 300)
```

# Further reading

Another potential way to evaluate shipping traffic is to obtain information from vessel traffic density such as in [Womersley 2022](https://www.pnas.org/doi/10.1073/pnas.2117440119). In this paper, gridded products were **purchased** from Exact Earth (https://www.exactearth.com) for the years of 2011 to 2014 at 0.25° × 0.25° grid cell resolution.

[Spire maritime](https://spire.com/maritime/?utm_campaign=maritime_2022_exactearth_redirect&utm_source=exactearth&utm_medium=website&utm_content=homepage) also provides this information if [purchased](https://spire.com/maritime/use-cases/environmental-protection/).

[Marine traffic](marinetraffic.com) is also an interesting page to visit.

For the legend inspiration comes from [Global Marine Traffic](https://globalmaritimetraffic.org/gmtds.html).

















               