---
title: "shipping lines and shipping traffic"
description: |
 Create a map including shipping lines or traffic information.
author:
  - name: Miriam Lerma
    url: {}
date: 2025-06-12
output:
  distill::distill_article:
    self_contained: false
---

# Intro

# Shipping lanes

## Data

The data set contains the spatial designations of the maritime spatial plan published on 1 September 2021 as an additional information service. The planning scale is 1: 400,000. The binding documents are available at [BSH](https://www.bsh.de/DE/THEMEN/Offshore/Meeresraumplanung/Raumordnungsplan_2021/raumordnungsplan-2021_node.html)

To download: go to [GeoSeaPortal](https://www.geoseaportal.de/atomfeeds/Raumordnungsplan_AWZ_en.xml#download=1) The zip contains several shapefiles including Shipping

Alternatively, the data shapefile is available in the package GermanNorthSea.

```{r}
library(GermanNorthSea)
```

```{r}
German_Shipping<-GermanNorthSea::German_Shipping
```

Use **geom_sf** to view the German Shipping Lines

```{r}
library(tidyverse)
```

```{r, fig.height=5, fig.width=5}
ggplot2::ggplot()+ 
  ggplot2::geom_sf(data = German_Shipping, colour = "blue", fill= 'blue',alpha=0.2, lwd = 1)+
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))
```
    
To have reference of the area, include the land, coast and natural protected areas. These shapefiles are also available in the packagfe **GermanNorthSea**. 

```{r, fig.height=5, fig.width=5}
ggplot(German_Shipping)+ 
  geom_sf(data = German_EEZ,colour = "red", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast,colour = "red", fill= '#bde0fe',alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = 'black', fill = '#ffffbe')+
  geom_sf(data = German_natura, colour = "#3d6d22", fill= '#3d6d22',alpha=0.2, lwd = 0.5)+
  geom_sf(data = German_Shipping, aes(fill= Type), colour = "grey",alpha=0.9, size=0.5)+
  scale_fill_manual(values = c("#E7F7C6", "#77D7B3", "#00BEBE"))+
  coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  theme(legend.position = 'bottom')
```

Change CRS to 4326 if preferred, using the function **st_transform** from the package **sf**.

```{r}
library(sf)
```

```{r}
German_EEZ<-st_transform(German_EEZ, 4326)
German_coast<-st_transform(German_coast, 4326)
German_land<-st_transform(German_land, 4326)
German_Shipping<-st_transform(German_Shipping, 4326)
```

## Definitions

In the national MSP plans, countries have designated certain areas as priority areas for shipping. This means that shipping is prioritized in these areas before other types of activities when planning for use and management of the area. However, other activities may take place as long as they do not interfere with shipping. Countries surrounding the North Sea have applied different criteria for designating priority areas for shipping. Below is a description of the basis on which the different countries designation priority shipping areas, as well as maps that shows where these areas are located. [Read more](https://northsearegion.eu/northsee/s-hipping/nationally-designated-priority-areas-for-shipping/)

Germany uses three types of zoning for the implementation of its spatial plans. These include “priority areas” where one use (for example shipping, pipelines, etc.) is granted priority over all other spatially significant uses; “reservation areas” where one use is given special consideration in a comparative evaluation with other spatially significant planning tasks, measures and projects; and “marine protected areas” where measures are applicable for the reduction of impacts (e.g., through pollution) on the marine environment.

```{r,  fig.height=5, fig.width=5}
ggplot()+
  geom_sf(data = German_EEZ, colour = "white", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "white", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = '#858585', fill = '#C8C8C8')+
  
  geom_sf(data = German_Shipping,
          aes(fill= Type), color='transparent',alpha=0.7)+
  
  theme_bw()+
  theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        panel.background = element_rect(fill = '#009BC4'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text=element_text(color="white",size=12),
        legend.title=element_text(color="white",size=12),
        axis.text.y.right=element_blank(),
        axis.ticks.y.right=element_blank(),
        axis.text.x.top=element_blank(),
        axis.ticks.x.top=element_blank(),
        panel.border = element_rect(colour = '#005AAA', fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
   scale_fill_manual(name='Shipping lanes',
                     values = c("#E7F7C6", "#77D7B3", "#00BEBE"))+
  theme(legend.position = c(0.30,0.30))
```

### Priority area

The lines on the map are designated priority shipping. This implies that activities that have the potential to interfere with shipping cannot occur in these areas, but does not mean that these areas are exclusive to shipping.

The designation of these corridors is mainly based on AIS records. The spatial distribution of corridors often highlights the importance of good accessibility to the ports. Consideration was given to the economic implications for the ports if ships divert from their traditional routes as a result of obstructions such as offshore wind farms.

```{r,  fig.height=5, fig.width=5}
ggplot()+
  geom_sf(data = German_EEZ, colour = "white", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "white", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = '#858585', fill = '#C8C8C8')+
  
  geom_sf(data = subset(German_Shipping,German_Shipping$Type=='priority area'),
                        aes(fill= Type), color='transparent',alpha=0.7)+
  
  theme_bw()+
  theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        panel.background = element_rect(fill = '#009BC4'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text=element_text(color="white",size=12),
        legend.title=element_text(color="white",size=12),
        axis.text.y.right=element_blank(),
        axis.ticks.y.right=element_blank(),
        axis.text.x.top=element_blank(),
        axis.ticks.x.top=element_blank(),
        panel.border = element_rect(colour = '#005AAA', fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
   scale_fill_manual(name='Shipping lanes',
                     values = c("#E7F7C6"))+
  theme(legend.position = c(0.30,0.30))
```

### Temporary priority area

A temporary priority area is a zone temporarily designated to prioritize certain maritime activities due to specific needs or circumstances. For example, during periods of high traffic or special events, certain areas may be temporarily designated as priority zones to manage congestion and ensure the smooth flow of maritime operations.

```{r, fig.height=5, fig.width=5}
ggplot()+
  geom_sf(data = German_EEZ, colour = "white", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "white", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = '#858585', fill = '#C8C8C8')+
  
  geom_sf(data = subset(German_Shipping,German_Shipping$Type=='temporary priority area'),
          aes(fill= Type), color='transparent',alpha=0.7)+
  
  theme_bw()+
  theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        panel.background = element_rect(fill = '#009BC4'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text=element_text(color="white",size=12),
        legend.title=element_text(color="white",size=12),
        axis.text.y.right=element_blank(),
        axis.ticks.y.right=element_blank(),
        axis.text.x.top=element_blank(),
        axis.ticks.x.top=element_blank(),
        panel.border = element_rect(colour = '#005AAA', fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
   scale_fill_manual(name='Shipping lanes',
                     values = c("#77D7B3"))+
  theme(legend.position = c(0.30,0.30))
```

### Temporary reservation area

“Reservation areas” are where one use is given special consideration in a comparative evaluation with other spatially significant planning tasks.

A temporary reservation area refers to a space set aside for a specific purpose for a limited time. In the context of port operations, this could involve designating an area for vessels to wait before docking or for specific cargo handling needs. Such reservations help in managing port congestion and ensuring efficient operations.

```{r,  fig.height=5, fig.width=5}
ggplot()+
  geom_sf(data = German_EEZ, colour = "white", fill= NA,alpha=0.9, lwd = 0.5)+
  geom_sf(data = German_coast, colour = "white", fill= NA,alpha=0.9, lwd = 0.5,linetype="dashed")+
  geom_sf(data = German_land, colour = '#858585', fill = '#C8C8C8')+
  
  geom_sf(data = subset(German_Shipping,German_Shipping$Type=='temporary reservation area'),
          aes(fill= Type), color='transparent',alpha=0.7)+
  
  theme_bw()+
  theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        panel.background = element_rect(fill = '#009BC4'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text=element_text(color="white",size=12),
        legend.title=element_text(color="white",size=12),
        axis.text.y.right=element_blank(),
        axis.ticks.y.right=element_blank(),
        axis.text.x.top=element_blank(),
        axis.ticks.x.top=element_blank(),
        panel.border = element_rect(colour = '#005AAA', fill=NA, size=1.5))+
  
  coord_sf(xlim = c(3,9), ylim = c(53,56),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
   scale_fill_manual(name='Shipping lanes',
                     values = c("#00BEBE"))+
  theme(legend.position = c(0.30,0.30))
```


# Vessel density

## Data

The Vessel Density maps in the EU are created since the 2019 by Cogea for the European Marine Observation and Data Network [EMODnet](https://emodnet.ec.europa.eu/geoviewer/#!/). The dataset is updated every year and is available for viewing and download on EMODnet Human Activities web portal (https://emodnet.ec.europa.eu/en/human-activities). The maps are based on AIS data yearly purchased from Collecte Localisation Satellites (CLS) and ORBCOMM. The maps, GeoTIFF format, show shipping density in 1x1km cells of a grid covering all EU waters and some neighbouring areas. Density is expressed as hours per square kilometre per month. The following ship types are available:0 Other, 1 Fishing, 2 Service, 3 Dredging or underwater ops, 4 Sailing, 5 Pleasure Craft, 6 High speed craft, 7 Tug and towing, 8 Passenger, 9 Cargo, 10 Tanker, 11 Military and Law Enforcement, 12 Unknown and All ship types. Data are available by month of year. Yearly averages are also available.

To download go to [EMODnet Map Viewer](https://emodnet.ec.europa.eu/geoviewer/) > Catalogue > EMODnet Human Activities > Vessel density > Monthly totals 2017-2023 > All Types > Click on the "i" icon > Select Download.

A zip file named "EMODnet_HA_Vessel_Density_all_2017-2023 would be downloaded. For the period of December 2023, the raster information is under **vesseldensity_all_20231201.tif**.

The original file includes information from a larger area, to focus on a specific, the function **crop** can be used. [Crop](https://search.r-project.org/CRAN/refmans/raster/html/crop.html) returns a geographic subset of an object as specified by an Extent object (or object from which an extent object can be extracted/created). If x is a Raster* object, the Extent is aligned to x. Areas included in y but outside the extent of x are ignored (see extend if you want a larger area). 


```{r, echo=TRUE, eval=FALSE, message=FALSE}
# Read original file 
your_folder<-paste0(here::here(),"/_posts/2025-06-12-shippingtraffic")
your_file<-paste0(your_folder,"/vesseldensity_all_20231201.tif")
ShippingTraffic<-terra::rast(your_file)
# Plot
ggplot2::ggplot() +
  tidyterra::geom_spatraster(data = ShippingTraffic) +
  NULL
# Crop
library(raster)
polygon_crop <- as(extent(3790000,4250000, 3350000,3680000), 'SpatialPolygons')
crs(polygon_crop) <- "+proj=longlat +datum=WGS84 +no_defs"
ShippingTraffic_crop<- crop(ShippingTraffic, polygon_crop)
# Export
writeRaster(ShippingTraffic_crop, "ShippingTraffic_crop.tif")
```

## Load

Use the package **terra** to use the function **rast**.  

```{r, message=FALSE}
library(terra)
```

The function **rast** reads your file as a formal class SpatRaster

```{r}
your_folder<-paste0(here::here(),"/_posts/2025-06-12-shippingtraffic")
your_file<-paste0(your_folder,"/ShippingTraffic_crop.tif")
```

```{r}
ShippingTraffic<-rast(your_file)
```

```{r, eval=FALSE}
crs(ShippingTraffic)
```

```{r, fig.height=5, fig.width=5}
ggplot2::ggplot() +
  tidyterra::geom_spatraster(data = ShippingTraffic) +
  NULL
```

## Crop and Mask

Select only values inside the EEZ

```{r}
German_EEZ<-GermanNorthSea::German_EEZ
polygon_crop <- German_EEZ
```

```{r, fig.width=5, fig.height=5}
ShippingTraffic_crop <- crop(ShippingTraffic, polygon_crop)
ShippingTraffic_mask <- mask(x= ShippingTraffic_crop, mask=polygon_crop)
plot(ShippingTraffic_mask)
```

## Data frame

```{r, message=FALSE}
library(tidyverse)
```

```{r}
ShippingTraffic_df <- as.data.frame(ShippingTraffic_mask, xy = TRUE)
beepr::beep(sound=1)
```

The function rename, allows to change the name of the column. The radiance information is in the third column.

```{r}
ShippingTraffic_df<-ShippingTraffic_df %>%
  rename(vessel_density=3)
```

Check if the density values are plausible.

```{r}
range(ShippingTraffic_df$vessel_density)
```

Values distribution

```{r, fig.height=5, fig.width=5}
hist(ShippingTraffic_df$vessel_density)
```


Classification

```{r}
ShippingTraffic_class<-ShippingTraffic_df %>%
 mutate(
    density_class = case_when(
      vessel_density <= 0  ~ "class_A_0",
      vessel_density > 0 &  vessel_density < 1 ~ "class_A_0_1",
      vessel_density > 1 &  vessel_density < 2 ~ "class_A_1_2",
      vessel_density > 2 &  vessel_density < 3 ~ "class_A_2_3",
      vessel_density > 3 &  vessel_density < 4 ~ "class_A_3_4",
      vessel_density > 4 &  vessel_density < 5 ~ "class_A_4_5",
      vessel_density > 5 &  vessel_density < 6 ~ "class_A_5_6",
      vessel_density > 6 &  vessel_density < 7 ~ "class_A_6_7",
      vessel_density > 7 &  vessel_density < 8 ~ "class_A_7_8",
      vessel_density > 8 &  vessel_density < 9 ~ "class_A_8_9",
      vessel_density > 9 &  vessel_density < 10 ~ "class_A_9_10",
      
      vessel_density >= 10 &  vessel_density < 100 ~ "class_B_10_100",
      vessel_density >= 100 &  vessel_density < 1000 ~ "class_C_100_1000",
      vessel_density >= 1000 &  vessel_density < 10000 ~ "class_D_1000_10000",
      vessel_density >= 10000 &  vessel_density < 100000 ~ "class_E_10000_100000",
      TRUE  ~ "unknown"
    ))
```

The majority of values are close to 0

```{r}
ShippingTraffic_class %>%
  group_by(density_class)%>%
  tally()
```


## Select area

Use the function **geom_spatraster** to plot the vessel density data.  

```{r}
library(tidyterra)
```

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask)
```

## Change fill

The largest differences are noticeable on the lowest densities.

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +
  scale_fill_gradient(low = "#00BEBE", 
                      high = "#E7F7C6", 
                      na.value = "grey93", 
                      limits = c(0,5),
                      breaks = c(0,1,2,3,4,5,6,7,8,9,10),
                      )+
  NULL
```


## Land

Add land for reference.

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradient(low = "#00BEBE", 
                      high = "#E7F7C6", 
                      na.value = "grey93", 
                      limits = c(0,5),
                      breaks = c(0,1,2,3,4,5,6,7,8,9,10),
                      )
```


## Theme

Change background colors

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
    scale_fill_gradient(low = "#00BEBE", 
                      high = "#E7F7C6", 
                      na.value = "transparent", 
                      limits = c(0,5),
                      breaks = c(0,1,2,3,4,5,6,7,8,9,10),
                      )+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.20,0.30),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))
```


# Shipping Lanes and Traffic

## Low traffic 

```{r}
color_K <-"#007f5f"
color_J <-"#2b9348"
color_I <-"#55a630"
color_H <-"#80b918"
color_G <-"#aacc00"
color_F <-"#bfd200"
color_E <-"#d4d700"
color_D <-"#dddf00"
color_C <-"#eeef20"
color_B <-"#ffff3f"
palette_yellow<-c(color_B,color_C,color_D,color_E,color_F,color_G,color_H,color_I,color_J,color_K)
```

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023',
                      na.value = "transparent", 
                      colours = palette_yellow,
                      limits = c(0.00001,10),
                      breaks = c(2,4,6,8,10),
                      )+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.20,0.30),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))
```


## Medium traffic 

```{r}
color_I <-"#081c15"
color_H <-"#1b4332"
color_G <-"#2d6a4f"
color_F <-"#40916c"
color_E <-"#52b788"
color_D <-"#74c69d"
color_C <-"#95d5b2"
color_B <-"#b7e4c7"
color_A <-"#d8f3dc"
palette_green<-c(color_A,color_B,color_C,color_D,color_E,color_F,color_G,color_H,color_I)
```


```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023',
                      na.value = "transparent", 
                      colours = palette_green,
                      limits = c(5,100),
                      breaks = c(5,20,40,60,80,100),
                      )+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.20,0.30),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))
```


## High traffic 

```{r}
color_A <-"#e01e37"
color_B <-"#da1e37"
color_C <-"#c71f37"
color_D <-"#bd1f36"
color_E <-"#b21e35"
color_F <-"#a71e34"
color_G <-"#a11d33"
color_H <-"#85182a"
color_I <-"#6e1423"
color_J <-"#641220"
palette_red<-c(color_A,color_B,color_C,color_D,color_E,color_F,color_G,color_H,color_I,color_J)
```

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n (hrs per km2)',
                      na.value = "transparent", 
                      colours = palette_red,
                      limits = c(100,30000),
                      breaks = c(0,1000,10000,20000,30000),
                      )+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.20,0.30),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))
```

# Export

## Palette

```{r}
your_palette<-c("transparent",
                "#d8f3dc",
                "#b7e4c7",
                "#95d5b2",
                "#74c69d",
                "#52b788",
                "#aacc00",
                "#bfd200",
                "#d4d700",
                "#dddf00",
                "#eeef20",
                "#ffff3f",
                "#e01e37",
                "#da1e37",
                "#c71f37",
                "#bd1f36",
                "#b21e35",
                "#a71e34",
                "#a11d33",
                "#85182a",
                "#6e1423",
                "#641220")
```

## Legend

Use **scale_fill_gradientn** to adjust the palette

```{r, fig.height=5, fig.width=5}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(0,100,30000),
                      values = scales::rescale(c(0,1,2,2,3,5,10,100,1000,10000,30000)))+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.20,0.50),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))+
    theme(legend.key.height = unit(12, "mm"))
```

Use **rescale** to adjust the legend

```{r}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(
                                 0.05,
                                 0.1,
                                 0.2,
                                 0.5,
                                 2,
                                 5,
                                 10,
                                 20,
                                 100,
                                 1000),
                     
                      values = scales::rescale(c(0,
                                                 0.01,
                                                 0.05,
                                                 0.1,
                                                 0.2,
                                                 0.5,
                                                 2,
                                                 5,
                                                 10,
                                                 20,
                                                 100,
                                                 1000)), 
                      guide = "legend"
                      )+
   
  theme_void()+
    theme(legend.background = element_rect(colour = "transparent", fill = "transparent"),
        legend.position = c(0.15,0.50),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=1.5),
        legend.text=element_text(color='black',size=12),
        legend.title=element_text(color='black',size=16))+
    theme(legend.key.height = unit(8, "mm"))
```

Use **labels** to make the legend more legible

```{r}
ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(
                                 0.05,
                                 0.1,
                                 0.2,
                                 0.5,
                                 2,
                                 5,
                                 10,
                                 20,
                                 100,
                                 1000),
                      values = scales::rescale(c(0,
                                                 0.01,
                                                 0.05,
                                                 0.1,
                                                 0.2,
                                                 0.5,
                                                 2,
                                                 5,
                                                 10,
                                                 20,
                                                 100,
                                                 1000)), 
                      labels = c(
                                 "0 - 0.05",
                                 "> 0.1",
                                 "> 0.2",
                                 "> 0.5",
                                 "> 2",
                                 "> 5",
                                 "> 10",
                                 "> 20",
                                 "> 100",
                                 "> 1,000"),
                      guide = "legend"
                      )+
   
  theme_void()+
    theme(
          panel.background = element_rect(fill = '#009BC4'),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
    theme(legend.position = c(0.15,0.50),
          legend.background = element_rect(colour = FALSE, fill = FALSE),
          legend.title=element_text(color='white',size=16),
          legend.text=element_text(color='white',size=12),
          legend.key = element_rect(colour = 'transparent', fill = 'transparent'),
          legend.key.height = unit(8, "mm"))
```

```{r, echo=FALSE, eval=FALSE}
ggsave(ggplot() +
  geom_spatraster(data = ShippingTraffic_mask) +

  geom_sf(data = German_EEZ, color='grey',fill='transparent',alpha=0.1, size = 1)+
  
  geom_sf(data = GermanNorthSea::German_land, colour = 'black', fill = '#e5e5e5')+
  
  #geom_sf(data = GermanNorthSea::German_Shipping, color='grey',fill='transparent',alpha=0.1, linetype = "dotted")+
  
  ggplot2::coord_sf(xlim = c(3790000,4250000), ylim = c(3350000,3680000),
                    label_axes = list(top = "E", left = "N", bottom = 'E', right='N'))+
  
  scale_fill_gradientn(name='Vessel density \n Dec 2023 \n hrs per km2',
                      na.value = "transparent", 
                      colours = your_palette,
                      limits = c(0,30000),
                      breaks = c(
                                 0.05,
                                 0.1,
                                 0.2,
                                 0.5,
                                 2,
                                 5,
                                 10,
                                 20,
                                 100,
                                 1000),
                      values = scales::rescale(c(0,
                                                 0.01,
                                                 0.05,
                                                 0.1,
                                                 0.2,
                                                 0.5,
                                                 2,
                                                 5,
                                                 10,
                                                 20,
                                                 100,
                                                 1000)), 
                      labels = c(
                                 "0 - 0.05",
                                 "> 0.1",
                                 "> 0.2",
                                 "> 0.5",
                                 "> 2",
                                 "> 5",
                                 "> 10",
                                 "> 20",
                                 "> 100",
                                 "> 1,000"),
                      guide = "legend"
                      )+
   
  theme_void()+
    theme(
          panel.background = element_rect(fill = '#009BC4'),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_rect(colour = "black", fill=NA, size=1.5))+
  
    theme(legend.position = c(0.15,0.50),
          legend.background = element_rect(colour = FALSE, fill = FALSE),
          legend.title=element_text(color='white',size=16),
          legend.text=element_text(color='white',size=12),
          legend.key = element_rect(colour = 'transparent', fill = 'transparent'),
          legend.key.height = unit(8, "mm")),
  
       filename = paste0(your_folder,"/VesselDensity_GermanNorthSea.jpg"),
       width = 10, 
       height =10, 
       units = "in",
       dpi = 300)
```

# Further reading

Another potential way to evaluate shipping traffic is to obtain information from vessel traffic density such as in [Womersley 2022](https://www.pnas.org/doi/10.1073/pnas.2117440119). In this paper, gridded products were **purchased** from Exact Earth (https://www.exactearth.com) for the years of 2011 to 2014 at 0.25° × 0.25° grid cell resolution.

[Spire maritime](https://spire.com/maritime/?utm_campaign=maritime_2022_exactearth_redirect&utm_source=exactearth&utm_medium=website&utm_content=homepage) also provides this information if **purchased**(https://spire.com/maritime/use-cases/environmental-protection/).

[Marine traffic](marinetraffic.com)

For the legend use: [Global Marine Traffic](https://globalmaritimetraffic.org/gmtds.html)

Example in [publication](https://www.researchgate.net/publication/341642158_Potential_impact_of_wash_water_effluents_from_scrubbers_on_water_acidification_in_the_southern_North_Sea)


















               