---
title: "Subset shapefile"
description: |
  Extract a specific polygon from a shapefile and export it as new shapefile.
author:
  - name: Miriam Lerma
    url: {}
date: 02-04-2023
preview: blog20.jpg
categories:
  - QGIS
  - shapefiles
  - English
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    self_contained: false
---

# Intro

Three steps to subset a shapefile and export a new shapefile with the attributes you are interested in.

# 1. Load shapefile

Load the package **sf** to load your shapefile into R

```{r}
library(sf)
```

Load the package **here** to use your directory (in which folder is your shapefile)

```{r}
library(here)
```

```{r, echo=FALSE}
My_directory<-here('_posts/2023-02-04-subset-shapefile')
```

Load your shapefile, be careful with your directory

```{r}
Old_shapefile<- read_sf(paste0(My_directory,'/original_shapefile.shp'))
```

Check the class, theoretically it would show sf

```{r}
class(Old_shapefile)
```

Explore the contents

```{r, eval=FALSE}
str(Old_shapefile)
```

Check the values in the column that you are interested to subset, to show unique values use the argument **unique** and to show them in alphabetic order use **sort**

```{r, eval = FALSE}
sort(unique(Old_shapefile$comuna))
```

# 2. Filter

Load the package **tidyverse**

```{r, message = FALSE, echo=TRUE}
library(tidyverse)
```

Using the function **filter** you can subset your old new shapefile and get the new one

In the example we used the column *comuna* and the value we are interested is *La Higuera*, do not forget to replace those values

```{r}
New_shapefile<-Old_shapefile %>%
  filter(comuna=="La Higuera")
```

# 3. Export your new shapefile

Check the class of your new object

```{r}
class(New_shapefile)
```

Export to your selected directory

```{r, echo=TRUE, eval=FALSE}
st_write(New_shapefile, paste0(My_directory,'/New_shapefile.shp'))
```

# Read more

- [Geocomputation with R](https://geocompr.robinlovelace.net/attr.html)
