---
title: "Create a buffer"
description: |
  This post is about how to create a spatial buffer of 1 km around a point.
author:
  - name: Miriam Lerma
    url: {}
date: 2024-03-11
preview: blog32a.jpg
output:
  distill::distill_article:
    self_contained: false
categories:
  - R
  - ggplot2
  - English
  - Y2024
---


# Intro

In this blog, a buffer subseting environmental data would be created and the average values within the buffer calculated.

# 1. Download data

Load the package **rerddap**

```{r}
library("rerddap")
```

```{r}
SST_info <- info('erdMWsstd1day_LonPM180')
```

To subset the data select the coordinates of a smaller area. 

```{r}
lonmin<--111
lonmax<--109
latmin<-23
latmax<-27
```

To download, provide the parameters such as the server, the coordinates and the time frame. It takes some time to download.

```{r}
SST_values<-griddap(SST_info, 
                    latitude= c(latmin, latmax), longitude = c(lonmin, lonmax), 
                    time = c('2023-03-01T00:00:00Z','2023-03-30T00:00:00Z'), 
                    fields = 'sst')
```
To acces the data as data frame, extract the data. 

```{r}
SST_df<-SST_values$data
```

Using functions from the package tidyverse, remove all NAs.

```{r, warning=FALSE}
library(tidyverse)
```

```{r}
SST_dfclean<-SST_df %>%
  filter(sst!='NaN')
```

## Map 

Check data downloaded using ggplot. 

```{r}
library(ggplot2)
```

First, download data from land from the package **rworldmap**.

```{r}
library(rworldmap)
```

The function *getMap()* loads the map in your environment.

```{r}
world_map <- getMap()
```

To plot using ggplot a data frame is recommended. Use the function *fortify* to be able to plot the map using ggplot.

```{r}
world_points <- fortify(world_map)
world_points$region <- world_points$id
world_df <-world_points[,c("long","lat","group", "region")]
```

```{r, fig.height=4}
ggplot() +
 geom_raster(data=SST_dfclean,aes(x=lon, y=lat, fill = sst))+ scale_fill_viridis_c(option = "H")+
 geom_polygon(data = world_df, aes(x = long, y = lat, group = group), colour = '#403d39', fill = "#e5e5e5") +
 coord_sf(xlim = c(-109.7,-109.1),ylim = c(24.6,25.2))
```



# 2. Create buffer

Load package **sp**.

```{r}
library(sp)
```

Give coordinates from point. 

```{r}
This_point<-data.frame(Longitude=-109.37728617302638,Latitude=24.92572600240793)
```

This function is a custom function, that is also included in my package **larus**.

```{r}
create_buffer<-function(central_point=central_point, buffer_km=buffer_km){
  central_spatial<- sp::SpatialPoints(cbind(central_point$Longitude,central_point$Latitude)) 
  sp::proj4string(central_spatial)= sp::CRS("+init=epsg:4326") 
  central_spatial <- sp::spTransform(central_spatial, sp::CRS("+init=epsg:4326"))
  buffer_dist<-buffer_km/100
  central_buffer<-rgeos::gBuffer(central_spatial, width=1*buffer_dist)
  return(central_buffer)
  }
```

The arguments needed are the point, and the kilometers.

```{r}
This_buffer<-create_buffer(central_point=This_point,buffer_km=20)
```
Then, I use the projection to convert it to SpatialPolygons, and ready for the next step.

```{r}
polygon_proj <- proj4string(This_buffer)
```

```{r}
class(This_buffer)
```

## Map

Make sure that they are in the same coordinate system. 

```{r, fig.height=4}
ggplot() +
 geom_raster(data=SST_dfclean,aes(x=lon, y=lat, fill = sst))+ scale_fill_viridis_c(option = "H")+
 geom_polygon(data = world_df, aes(x = long, y = lat, group = group), colour = '#403d39', fill = "#e5e5e5") +
 coord_sf(xlim = c(-109.7,-109.1),ylim = c(24.6,25.2))+
 geom_polygon(data=This_buffer,aes(x = long, y = lat, group = group),colour='black', fill='transparent', linetype='dashed')+
 geom_point(data=This_point, aes(x=Longitude, y=Latitude), colour = "black", size = 4)
```



# 3. Extract information

Now that we have our buffer and our SST, lets convert the SST from data frame to SpatialPointsDataFrame. 

```{r}
SST_sp <- SST_dfclean
sp::coordinates(SST_sp) <- ~lon + lat
sp::proj4string(SST_sp) = sp::CRS("+init=epsg:4326")
```

Use the function **over** from the package **sp**. 

```{r}
SST_buffer<- over(SST_sp,This_buffer)
```

Now we have information on those values that fall inside the buffer. Add this information to the data frame. 

```{r}
SST_dfclean$sst_inside <- as.numeric(SST_buffer)
```

Now, subset to only keep the data inside the buffer. 

```{r}
SST_inside<-SST_dfclean %>%
  filter(sst_inside==1)%>%
  drop_na(sst)
```


## Subset

```{r, fig.height=4}
ggplot() +
  geom_raster(data=SST_inside,aes(x=lon, y=lat, fill = sst))+ scale_fill_viridis_c(option = "H")+
  geom_polygon(data = world_df, aes(x = long, y = lat, group = group), colour = '#403d39', fill = "#e5e5e5") +
  coord_sf(xlim = c(-109.7,-109.1),ylim = c(24.6,25.2))+
  geom_polygon(data=This_buffer,aes(x = long, y = lat, group = group),colour='black', fill='transparent', linetype='dashed')+
  geom_point(data=This_point, aes(x=Longitude, y=Latitude), colour = "black", size = 4)+
  theme(panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill='transparent'))
```

```{r, echo=FALSE, eval=FALSE}
library(here)
here()
ggsave(ggplot() +
  geom_raster(data=SST_inside,aes(x=lon, y=lat, fill = sst))+ scale_fill_viridis_c(option = "H")+
  geom_polygon(data = world_df, aes(x = long, y = lat, group = group), colour = '#403d39', fill = "#e5e5e5") +
  coord_sf(xlim = c(-109.7,-109.1),ylim = c(24.6,25.2))+
  geom_polygon(data=This_buffer,aes(x = long, y = lat, group = group),colour='black', fill='transparent', linetype='dashed')+
  geom_point(data=This_point, aes(x=Longitude, y=Latitude), colour = "black", size = 4)+
  theme(panel.background = element_blank(),
        panel.border = element_rect(colour = "black", fill='transparent')), 
       filename = paste0(here(),"/_posts/2024-04-11-buffer/blog32.jpg"),
       width = 6, 
       height = 4, 
       units = "in",
       dpi = 300)
```

End of document